---
title: "icvs_2015_prep_report"
output:
  html_document:
    df_print: paged
    number_sections: true
    toc: true
---

```{r source script 1 & data, warning=F, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(haven) # haven for reading .sav file
library(janitor)
library(stats)
library(tidyverse)
library(dplyr) # for glimpse and filter functions
library(data.table)
library(knitr)
library(readxl) # read xls file

source(here::here("scripts", "s2_01_icvs_scripts.r"), local = knitr::knit_global())

#sjPlot::view_df(icvs_data)
#names_and_labels <- icvs_data %>%
#  surveytoolbox::varl_tb()

#names_and_labels <- icvs_data %>%
#surveytoolbox::extract_vallab()
#raw_classes <- as.character(lapply(lapply(icvs_data, class),tail,1))

icvs_data <- read_sav("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/ICVS2015_2b.sav")  # Reading data

# names(icvs_data)
# 
# # urban or rural
# glimpse(icvs_data$i004)
# 
# # main city
# glimpse(icvs_data$i004a)
# 
# glimpse(icvs_data$i005b)
# summary(as.factor(icvs_data$i005b))
```


```{r janitor, message=FALSE, echo=FALSE}
raw_classes <- sapply(sapply(icvs_data, class),tail,1)

icvs_data <- icvs_data %>% 
  mutate_all(as_factor) %>%   # convert all variables from chr+lbl to fct type          
  set_names(icvs_data %>%       # convert variable names to informative labels
              sjlabelled::get_label() %>% # pull labels from original tibble
              enframe() %>%               # convert list of column names to a new tibble
              na_if("") %>%               # if variable label is empty string, convert to NA
              mutate(value = coalesce(value, name)) %>%  # fill in NA with original if missing
              pull(value)) %>%            # extract new variable name into a character vector
  janitor::clean_names()      # clean names to be all lowercase, replacing spaces with "_"     

```



```{r unlabeled,message=FALSE, echo=FALSE}
# i005b identifies the national surveys conducted

summary(icvs_data$i002b)
summary(icvs_data$sweep)

# i004 appears to be sweep information
```



The ICVS provides more specific regional data compared to the LIS or World Bank, distinguishing data collection across England and Wales, Ireland, Northern Ireland, and Scotland (in addition to the United Kingdom).



# Creating a variable for countries 

Since my analysis is multilivel, with countries as clusters, identifying countries in a single variable is imperative.

```{r nation-scan}
summary(icvs_data$national_surveys)
```

A lot of missing values for national_surveys

```{r main_city}
summary(icvs_data$main_city)
```

still a lot of NAs, but gives values for missing data like Turkey

```{r first country creation}
# extract characters between parentheses from main_city
icvs_data$country <- gsub("(?<=\\()[^()]*(?=\\))(*SKIP)(*F)|.", "", icvs_data$main_city, perl=T)

summary(as.factor(icvs_data$country))
# view unique countries
# unique(icvs_data$country)
```

Lots of blanks and NAs. Maybe we can supplement this with the national_surveys variable

```{r missingness}
# change blank to NA
icvs_data$country <- gsub("^$|^ $", NA, icvs_data$country)

# for missing country variables, import national survey values
icvs_data$country[is.na(icvs_data$country)] <- as.character(icvs_data$national_surveys[is.na(icvs_data$country)])

# view after national survey
summary(as.factor(icvs_data$country))

icvs_data$country <- as.factor(icvs_data$country)
```

gets rid of under half of NA values

```{r does this work, message=FALSE, echo=FALSE}
# I don't think this does anything
# icvs_data$country[is.na(icvs_data$country)] <- icvs_data$national_surveys[is.na(icvs_data$country)]
# 
# summary(as.factor(icvs_data$country))
```

Looking at the 2005.sav, I know Spain should be visible by now, fortunately, the regional_survey variable seems uniquely created to handle this

```{r regional spain}
summary(icvs_data$regional_survey)

# filling spain into country variable from regional_survey responses
icvs_data$country[is.na(icvs_data$country)] <- gsub("(?<=\\()[^()]*(?=\\))(*SKIP)(*F)|.", "", icvs_data$regional_survey[is.na(icvs_data$country)], perl=T)
```

```{r post-spain}
summary(icvs_data$country)
```

Spain is now in. Since I'm just using 2004-2006+ 2005 ICS, maybe I can just stop here. If there are no more missing values in my sweeps, I can just exclude the rest and move on

```{r sweep country missingness}
icvs_data$sweep[is.na(icvs_data$country)] %>%
  summary()

```

no luck

```{r nation-identifying}
icvs_data$country <- as.factor(icvs_data$country)


icvs_data$country<- recode(icvs_data$country, "SAR China"= "Hong Kong",
                           "USA" = "United States", "RSA" = "South Africa")

data_2005 <- icvs_data %>%
  filter(sweep == "5th sweep -2004-2006" | sweep == "EU ICS 2005") 

summary(data_2005$country)[summary(data_2005$country) >0]
```

6020 NAs

From 2005.sav file:

math checks out for united kingdom (889+1004+2002+2010+2004 = 7909)

missing cases for peru (7011 vs 1030 = 3289), argentina (2529 vs 5260= 2731);  2731+3289 = 6020. Accounts for all NAs


## Peru and argentina

This interactive table of a subset of rows suggests that these observations are lacking country identification. They have latin america for global region, but nothing more specific 

A combination of the region variable and questionnaire_that_was_used seems to hold some kind of promise. Reliably, the country-specific surveys have region information, bat FtoF 2000 does not.

```{r look at data}
set.seed(12345) 

data_2005[is.na(data_2005$country),][sample(1:nrow(data_2005[is.na(data_2005$country),]), 30), ]

```


Another perspective is to look at patterns of response country X survey type

```{r questionnaire missingness, message=FALSE}
setnames(data_2005, old=c("questionnaire_that_was_used","questionnaire_was_based_on"), new=c("questionnaire_used","questionnaire_based_on"), skip_absent=TRUE)


questionnaire_screen <- data_2005 %>% group_by(questionnaire_used) %>% 
  dplyr::count(country)

# view number of responses for each country*sweep
questionnaire_screen_c <- reshape::cast(questionnaire_screen,country~questionnaire_used)

# compute total number of participants
questionnaire_screen_c$total <- rowSums(questionnaire_screen_c[,2:ncol(questionnaire_screen_c)], na.rm=TRUE)

questionnaire_screen_c$total[questionnaire_screen_c$total == 0] <- NA

questionnaire_screen_c

```

Looking at the 2005.sav, Argentina got the FtoF 2000, and Peru got Country specific

```{r argentina and peru}
data_2005$country <- ifelse(is.na(data_2005$country),ifelse(data_2005$questionnaire_used == "FtoF 2000", "Argentina",ifelse(data_2005$questionnaire_used == "Country specific", "Peru",NA)),as.character(data_2005$country))

data_2005$country <- as.factor(data_2005$country)


summary(data_2005$country)
```

This treatment gives the same values as 2005.sav



## Uk transformations

```{r non adjusted countries}
summary(data_2005$country)[c("United Kingdom","England & Wales", "England", "Northern Ireland", "Scotland", "Ireland")]
```


It would be preferable to retain data granularity across the United Kingdom. Firstly, it would keep the number of clusters closer 30 countries, and secondly, the regions likely have important economic and cultural differences. For instance, Northern Ireland has a smaller gini compared to the broader UK (.33 vs. .39 in 2013; https://www.nicva.org/resource/economic-inequality-in-northern-ireland).

Unfortunately, smaller regions seem less likely to systematically collect data. For instance, some critical variables such as income inequality are not available either in the LIS, World Bank, nor government-level data in countries such as Northern Ireland or Scotland. Since this study seeks to integrate individual and nation-level observations, members of the United Kingdom were combined under the single country of the UK.

```{r adjusting countries}
# substituting specific british isles regions with UK
data_2005$country[data_2005$country %in% c("England & Wales", "England", "Northern Ireland", "Scotland")] <- "United Kingdom"
```


```{r total adjusted countries}
summary(data_2005$country)[c("United Kingdom","England & Wales", "England", "Northern Ireland", "Scotland", "Ireland")]
sum(summary(data_2005$country)>0)
```

Notably, Ireland is not part of the UK, so has better data coverage and is left separately for this study


# data_2015

```{r 2015-creation}
data_2015 <- icvs_data %>%
  filter(sweep == "Carribean 2014" |sweep == "Asia 2015 onwards")

unique(data_2015$country)
```

```{r 2010-creation}
data_2010 <- icvs_data %>%
  filter(sweep == "Carribean 2014" |sweep == "Asia 2015 onwards" | sweep == "6th sweep -2010")

unique(data_2010$country)
```

at most, the recent data only gives us `r length(unique(data_2010$country))` countries
