---
title: "icvs_2015_prep_report"
output:
  html_document:
    df_print: paged
    number_sections: true
    toc: true
---

```{r source script 1 & data, warning=F, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(haven) # haven for reading .sav file
library(janitor)
library(stats)
library(tidyverse)
library(dplyr) # for glimpse and filter functions
library(data.table)
library(knitr)
library(readxl) # read xls file

#sjPlot::view_df(icvs_data)
#names_and_labels <- icvs_data %>%
#  surveytoolbox::varl_tb()

#names_and_labels <- icvs_data %>%
#surveytoolbox::extract_vallab()
#raw_classes <- as.character(lapply(lapply(icvs_data, class),tail,1))

icvs_data <- read_sav("C:/Users/dalla/Google Drive/data_files/icvs_pwt_swiid/data/ICVS2015_2b.sav")  # Reading data

names(icvs_data)

# urban or rural
glimpse(icvs_data$i004)

# main city
glimpse(icvs_data$i004a)

glimpse(icvs_data$i005b)
summary(as.factor(icvs_data$i005b))
```

i005b identifies the national surveys conducted

The International Crime Victims Survey (ICVS) is a great dataset. Big, easy enough to access, and it's well documented. The identifiers and labels are available in the dataset, as well as  an accompanying codebook updated to the 2005 sweep of the ICVS. There is also a version of the questionnaire that participants see, which has been helpful in interpreting the items in plain language, in addition to clearing up some inconsistencies in the codebook.
```{r janitor, message=FALSE, echo=FALSE}
raw_classes <- sapply(sapply(icvs_data, class),tail,1)

icvs_data <- icvs_data %>% 
  mutate_all(as_factor) %>%   # convert all variables from chr+lbl to fct type          
  set_names(icvs_data %>%       # convert variable names to informative labels
              sjlabelled::get_label() %>% # pull labels from original tibble
              enframe() %>%               # convert list of column names to a new tibble
              na_if("") %>%               # if variable label is empty string, convert to NA
              mutate(value = coalesce(value, name)) %>%  # fill in NA with original if missing
              pull(value)) %>%            # extract new variable name into a character vector
  janitor::clean_names()      # clean names to be all lowercase, replacing spaces with "_"     

```



```{r unlabeled}
summary(icvs_data$i002b)
summary(icvs_data$sweep)
```

i004 appears to be sweep information

# Creating a variable for countries 

Since my analysis is multilivel, with countries as clusters, identifying countries in a single variable is imperative.

```{r nation-scan}
summary(icvs_data$national_surveys)
```

A lot of missing values for national_surveys

```{r main_city}
summary(icvs_data$main_city)
```

still a lot of NAs, but gives values for missing data like Turkey

```{r first country creation}
# extract characters between parentheses from main_city
icvs_data$country <- gsub("(?<=\\()[^()]*(?=\\))(*SKIP)(*F)|.", "", icvs_data$main_city, perl=T)

summary(as.factor(icvs_data$country))
# view unique countries
# unique(icvs_data$country)
```

Lots of blanks and NAs. Maybe we can supplement this with the national_surveys variable

```{r missingness}
# change blank to NA
icvs_data$country <- gsub("^$|^ $", NA, icvs_data$country)

# for missing country variables, import national survey values
icvs_data$country[is.na(icvs_data$country)] <- as.character(icvs_data$national_surveys[is.na(icvs_data$country)])

# view after national survey
summary(as.factor(icvs_data$country))

icvs_data$country <- as.factor(icvs_data$country)
```

gets rid of NA values

```{r does this work, message=FALSE, echo=FALSE}
# I don't think this does anything
# icvs_data$country[is.na(icvs_data$country)] <- icvs_data$national_surveys[is.na(icvs_data$country)]
# 
# summary(as.factor(icvs_data$country))
```

Looking at the 2005.sav, I know Spain should be visible by now, fortunately, the regional_survey variable seems uniquely created to handle this

```{r regional spain}
summary(icvs_data$regional_survey)

# filling spain into country variable from regional_survey responses
icvs_data$country[is.na(icvs_data$country)] <- gsub("(?<=\\()[^()]*(?=\\))(*SKIP)(*F)|.", "", icvs_data$regional_survey[is.na(icvs_data$country)], perl=T)
```

```{r post-spain}
summary(icvs_data$country)
```

Spain is now in. Since I'm just using 2004-2006+ 2005 ICS, maybe I can just stop here. If there are no more missing values in my sweeps, I can just exclude the rest and move on

```{r sweep country missingness}
icvs_data$sweep[is.na(icvs_data$country)] %>%
  summary()

```

no luck

```{r nation-identifying}
icvs_data$country <- as.factor(icvs_data$country)

data_2005 <- icvs_data %>%
  filter(sweep == "5th sweep -2004-2006" | sweep == "EU ICS 2005") 

summary(data_2005$country)[summary(data_2005$country) >0]
```

6020 NAs

From 2005.sav file:

math checks out for united kingdom (889+1004+2002+2010+2004 = 7909)

missing cases for peru (7011 vs 1030 = 3289), argentina (2529 vs 5260= 2731);  2731+3289 = 6020. Accounts for all NAs


## Peru and argentina

This interactive table of a subset of rows suggests that these observations are lacking country identification. They have latin america for global region, but nothing more specific 

A combination of the region variable and questionnaire_that_was_used seems to hold some kind of promise. Reliably, the country-specific surveys have region information, bat FtoF 2000 does not.

```{r look at data}
set.seed(12345) 

data_2005[is.na(data_2005$country),][sample(1:nrow(data_2005[is.na(data_2005$country),]), 30), ]

```


```{r questionnaire missingness}
setnames(data_2005, old=c("questionnaire_that_was_used","questionnaire_was_based_on"), new=c("questionnaire_used","questionnaire_based_on"), skip_absent=TRUE)


questionnaire_screen <- data_2005 %>% group_by(questionnaire_used) %>% 
  dplyr::count(country)

# view number of responses for each country*sweep
questionnaire_screen_c <- cast(questionnaire_screen,country~questionnaire_used)

# compute total number of participants
questionnaire_screen_c$total <- rowSums(questionnaire_screen_c[,2:ncol(questionnaire_screen_c)], na.rm=TRUE)

questionnaire_screen_c$total[questionnaire_screen_c$total == 0] <- NA

questionnaire_screen_c

```

Looking at the 2005.sav, Argentina got the FtoF 2000, and Peru got Country specific

```{r argentina and peru}
if (data_2005$questionnaire_used == "FtoF 2000" && is.na(data_2005$country)){
  data_2005$country <- "Argentina"
}

if (data_2005$questionnaire_used == "Country specific" && is.na(data_2005$country)){
  data_2005$country <- "Peru"
} 
 
summary(data_2005$country)[summary(data_2005$country) >0]
```
This treatment gives the same values as 2005.sav


# Other descriptive stuff


```{r ncols and nrow, warning=F, message=FALSE, include = TRUE}
nrow(icvs_data)
ncol(icvs_data)
```

As we can see, very large, `r ncol(icvs_data)` variables over `r nrow(icvs_data)` respondents

<br>



```{r country info}
unique(icvs_data$country)
length(unique(icvs_data$country))
```


<br>
We see above, that there are 84 countries in total across the entire ICVS. 


# **At a Glance**

```{r source scripts, warning=F, message=FALSE, echo=FALSE}
 source("C:/Users/dalla/Google Drive/project_files/icvs_inequality/scripts/icvs_scripts.r", local = knitr::knit_global())

#setnames(icvs_data3, old=c("i002a","i002b"), new=c("sweep_year", "sweep_num"), skip_absent=TRUE)

#attach names and classes to cleaned dataset

#library(hfunk)
#changeClass(icvs_data2, raw_classes)

```

The ICVS provides more specific regional data compared to the LIS or World Bank, distinguishing data collection across England and Wales, Ireland, Northern Ireland, and Scotland (in addition to the United Kingdom).

```{r non adjusted countries}
summary(icvs_data$country)[c("United Kingdom","England & Wales", "England", "Northern Ireland", "Scotland", "Ireland")]
```


It would be preferable to retain data granularity across the United Kingdom. Firstly, it would keep the number of clusters closer 30 countries, and secondly, the regions likely have important economic and cultural differences. For instance, Northern Ireland has a smaller gini compared to the broader UK (.33 vs. .39 in 2013; https://www.nicva.org/resource/economic-inequality-in-northern-ireland).

Unfortunately, smaller regions seem less likely to systematically collect data. For instance, some critical variables such as income inequality are not available either in the LIS, World Bank, nor government-level data in countries such as Northern Ireland or Scotland. Since this study seeks to integrate individual and nation-level observations, members of the United Kingdom were combined under the single country of the UK.

```{r adjusting countries, echo = FALSE}
# substituting specific british isles regions with UK
icvs_data$country[icvs_data$country %in% c("England & Wales", "England", "Northern Ireland", "Scotland")] <- "United Kingdom"

icvs_data$country<- recode(icvs_data$country, "SAR China"= "Hong Kong",
                           "USA" = "United States", "RSA" = "South Africa")
```


```{r total adjusted countries}
summary(icvs_data$country)[c("United Kingdom","England & Wales", "England", "Northern Ireland", "Scotland", "Ireland")]
sum(summary(icvs_data$country)>0)
```

Notably, Ireland is not part of the UK, so has better data coverage and is left separately for this study


# **Data wrangling**

There are many different years and versions of the survey; not every case, sweep, or questionnaire will yield the data we're looking for.

If we're doing international comparisons, the most important first consideration is the distribution of responses by sweep year. We need to have cases be (roughly) comparable based on the year. For example, it would be unfair to compare norway in 2005 to lithuania in 1989.

```{r more country info, warning=F, message=FALSE, echo=FALSE}
# kable_summary(icvs_data$sweep)
```


We can see that there are five major sweeps of data collection, each potentially spanning a couple years. There is also a specific EU sweep, overlapping the 5th sweep.

Next, it's worthwhile seeing what kinds of questionnaires people get, in case there are any meaningful differences


```{r questionnaires, warning=F, message=FALSE}
# kable_summary(icvs_data$questionnaire_used)
```

We can see  that *there are more than a dozen different surveys distributed as part of this project*: computer-assisted telephone interviewing (CATI), face to face, and something "country specific." I know with some questionnaires (possibly later, and outside this dataset) are done solely with computers, but those had very poor response rates. Likewise, we can see that the study has been conducted over 15 years, better seen below.

A potentially important note is that 9833 cases have "0" coded as the questionnaire they received, which doesn't have a code to it. Likewise, 16412 cases are NA. This means  about *25,000 participants were given an unknown questionnaire.* More on this later


<br>

This is important because the questions they receive, and the time they receive the questions, can change. There are  variables that may help us clear this up, but we need to keep this uncertainty in our consideration, because as explored later on, **this study often uses NA values in the place of 0** 

<br>

The codebook proves to be a useful, but incomplete, resource here. We can see which countries took part in each sweep, and what items were included (it even specifies the order of presentation!).

<br>

