---
title: "icvs_100_analyses"
author: "Dallas Novakowski"
date: "14/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(ordinal)
load("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/mod_joined.RData")

#used for dotwhisker plots
load("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/mod_joined1.RData")
```


```{r testmodel}
# m1 <- mod_joined %>% map(~ lmer(num_victim_5yr ~ gini_2004_6_cent + gdppc_2004_6_scale + 
#                      age_cent + employed + male + police_effective + income_quartile + (1 | country),
# data = .x))
```

```{r test tidymodel}
# m1_sims <- m1 %>%
#   map(. %>% arm::sim(n.sims = 1000)) %>%
#   map_df(. %>% slot("fixef") %>% as_tibble())
# 
# 
# # Generate a tidy dataframe of results (cf. broom::tidy())
# m1_tidy <- tibble(
#   term = names(m1_sims),
#   estimate = summarise_all(m1_sims, list(mean)) %>%
#     t() %>%
#     as.vector(),
#   std.error = m1_sims %>%
#     summarise_all(list(sd)) %>%
#     t() %>%
#     as.vector()
# )
# 
# 
# library(dotwhisker)
# # Plot the results
# m1_tidy %>%
#   by_2sd(mod_joined1) %>%
#   relabel_predictors(
#     c(
#       gini_2004_6_cent = "Income Inequality",
#       gdppc_2004_6_scale = "GDP/capita",
#       age_cent = "Age",
#       employed1 = "Employed",
#       male1 = "Male",
#       police_effective = "Perceived police effectiveness",
#       income_quartile = "Income quartile"
#     )
#   ) %>%
#   dwplot() +
#   theme_bw() +
#   geom_vline(xintercept = 0,
#              colour = "grey60",
#              linetype = "dashed") +
#   theme(legend.position = "none") +
#   labs(title = "Predicting crime victimization",
#        x = "Coefficient Estimate")

```

```{r m100_1}
m1_ord_null <- ordinal::clmm(ordered(num_victim_5yr_assault_winz) ~ 1 + (1 | country), data = mod_joined[[1]])

m1_ord <- ordinal::clmm(ordered(num_victim_5yr_assault_winz) ~  gdppc_2004_6_scale + gini_2004_6_cent + age_cent + employed + male + police_effective + income_quartile + (1 | country), data = mod_joined[[1]])

m1_ord_summ <- summary(m1_ord)

m1_ord$info$AIC
m1_ord$info$logLik
```


```{r ordinal model}
# takes about an hour
m100_ord <- mod_joined %>% map(~ ordinal::clmm(ordered(num_victim_5yr_assault_winz) ~  gdppc_2004_6_scale + gini_2004_6_cent + age_cent + employed + male + police_effective + income_quartile + (1 | country), data = .x))
# 
# m100_ord1 <- m100_ord[1:10]
# 
# save(m100_ord1, file = "C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord1.RData")
# 
# m100_ord2 <- m100_ord[11:20]
# 
# save(m100_ord2, file = "C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord2.RData")
# 
# m100_ord3 <- m100_ord[21:30]
# 
# save(m100_ord3, file = "C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord3.RData")
# 
# m100_ord4 <- m100_ord[31:40]
# 
# save(m100_ord4, file = "C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord4.RData")
# 
# m100_ord5 <- m100_ord[41:50]
# 
# save(m100_ord5, file = "C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord5.RData")
# 
# m100_ord6 <- m100_ord[51:60]
# 
# save(m100_ord6, file = "C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord6.RData")
# 
# m100_ord7 <- m100_ord[61:70]
# 
# save(m100_ord7, file = "C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord7.RData")
# 
# m100_ord8 <- m100_ord[71:80]
# 
# save(m100_ord8, file = "C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord8.RData")
# 
# m100_ord9 <- m100_ord[81:90]
# 
# save(m100_ord9, file = "C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord9.RData")
# 
# m100_ord10 <- m100_ord[91:100]
# 
# save(m100_ord10, file = "C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord10.RData")


# m100_ord2 <- mod_joined2 %>% map(~ ordinal::clmm(ordered(num_victim_5yr_winz) ~ gini_2004_6_cent + gdppc_2004_6_scale + age_cent + employed + male + police_effective + income_quartile + (1 | country), data = .x))
# 
# save(m100_ord2, file = "C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord2.RData")
```



```{r ordinal2 simulation}
list_100 <- list()

# load("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord1.RData")
# 


# load("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord2.RData")
# 
# load("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord3.RData")
# 
# load("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord4.RData")
# 
# load("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord5.RData")
# 
# load("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord6.RData")
# 
# load("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord7.RData")
# 
# load("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord8.RData")
# 
# load("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord9.RData")
# 
# load("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_ord10.RData")


library(stevemisc) # my toy R package with various helper functions
library(modelr) # for data_grid
library(kableExtra) # for pretty tables


# http://svmiller.com/blog/2020/04/summarizing-ordinal-models-with-simulation-multivariate-normal/
# extract coefficients for each model
m100_coef <- m100_ord %>%
  map(. %>% coef())

# extract
m100_vcov <- m100_ord %>%
  map(. %>% vcov()) 
  
m100_vcov <- m100_vcov %>%
  map(. %>% .[-nrow(.), -ncol(.)])

combined <- list()
combined <- mapply(data.frame, m100_coef, m100_vcov, SIMPLIFY=FALSE)

combined <- lapply(seq_along(combined), function(i) {
  colnames(combined[[i]])[1] <- "estimates"
  return(combined[[i]])
}) 

# m100_sims2 <- combined %>%
#   map(. %>% smvrnorm(1000, .$estimates, as.matrix(.[,-1])))

m100_sims <- list()

m100_sims <- lapply(seq_along(combined), function(i) {
 m100_sims[[i]] <- smvrnorm(1000, combined[[i]]$estimates, as.matrix(combined[[i]][,-1]))
  return(m100_sims[[i]])
}) 


save(m100_sims, file = "C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_sims.RData")
```


```{r m100 ord table}
load(file = "C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/m100_sims.RData")


m100_sims <- m100_sims %>%
  map_df(. %>% as_tibble())


m100_tidy <- tibble(
  term = names(m100_sims),
  estimate = summarise_all(m100_sims, list(mean)) %>%
    t() %>%
    as.vector(),
  std.error = m100_sims %>%
    summarise_all(list(sd)) %>%
    t() %>%
    as.vector()
)


library(dotwhisker)
# Plot the results
m100_tidy %>%
  by_2sd(mod_joined1) %>%
  relabel_predictors(
    c(
      gini_2004_6_cent = "Income Inequality",
      gdppc_2004_6_scale = "GDP/capita",
      age_cent = "Age",
      employed1 = "Employed",
      male1 = "Male",
      police_effective = "Perceived police effectiveness",
      income_quartile = "Income quartile"
    )
  ) %>%
  dwplot() +
  theme_bw() +
  geom_vline(xintercept = 0,
             colour = "grey60",
             linetype = "dashed") +
  theme(legend.position = "none") +
  labs(title = "Predicting crime victimization",
       x = "Coefficient Estimate")

# m100_sims2 <- smvrnorm(1000, combined[[1]][,1], as.matrix(combined[[1]][,-1])) 
  

# smvrnorm(1000, coefm100_ord1, vcovm100_ord1) %>% tbl_df() %>% # 1,000 sims, convert matrix to tbl
#   mutate(sim = seq(1:1000)) %>% # create simulation identifier
#   select(sim, everything())

```


```{r mord100}
m100_ord1 <- ordinal::clmm(ordered(num_victim_5yr_winz) ~ gini_2004_6_cent + gdppc_2004_6_scale + age_cent + employed + male + police_effective + income_quartile + (1 | country), data = mod_joined1)
```


```{r mord1 simulation}

simm100_ord1 <- smvrnorm(1000, coefm100_ord1, vcovm100_ord1) %>% tbl_df() %>% # 1,000 sims, convert matrix to tbl
  mutate(sim = seq(1:1000)) %>% # create simulation identifier
  select(sim, everything()) # make sim column first in the data
```


> Chapter 7 of Gelman and Hill (2007) for how the multivariate normal distribution is a novel way of simulating uncertainty regarding the model output from a generalized linear model. I’ll only note here that simulating values from a multivariate normal distribution of the ordinal model requires only the vector of regression coefficients and the variance-covariance matrix of the fitted model. 


```{r mw100-ord}
mw100_ord <- mod_joined %>% map(~ ordinal::clmm(ordered(num_victim_5yr_winz) ~ gini_2004_6_cent + gdppc_2004_6_scale + age_cent + employed + male + police_effective + income_quartile + (1 | country), data = .x, weights = individual_weight))

save(mw100_ord, file = "C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/mw100_ord.RData")
```


```{r view-convergence}
summary(mw100_ord[[1]])$coefficients[,"Std. Error"]

is.nan(summary(mw100_ord[[9]])$coefficients[,"Std. Error"]["gini_2004_6_cent"])

mw100_invalid <- list()

mw100_invalid <- lapply(seq_along(mw100_ord), function(i) {
 mw100_invalid[[i]] <- is.nan(summary(mw100_ord[[i]])$coefficients[,"Std. Error"]["gini_2004_6_cent"])
  return(mw100_invalid[[i]])
}) 

percent <- function(x, digits = 0, format = "f", ...) {      # Create user-defined function
  paste0(formatC(x * 100, format = format, digits = digits, ...), "%")
}

model_nonconvergence <- Reduce("+",mw100_invalid)

save(model_nonconvergence, file = here::here("output", "icvs_output.RData"))

```


```{r mw100 extraction}
# extract

mw100_coef2 <- mw100_ord %>%
  map(. %>% coef())


mw100_vcov2 <- mw100_ord %>%
  map(. %>% vcov()) 
  
combined <- list()
combined <- mapply(data.frame, mw100_coef2, mw100_vcov2, SIMPLIFY=FALSE)

combined <- lapply(seq_along(combined), function(i) {
  colnames(combined[[i]])[1] <- "estimates"
  return(combined[[i]])
}) 

# mw100_sims2 <- combined %>%
#   map(. %>% smvrnorm(1000, .$estimates, as.matrix(.[,-1])))

mw100_sims2 <- list()

mw100_sims2 <- lapply(seq_along(combined), function(i) {
 mw100_sims2[[i]] <- smvrnorm(1000, combined[[i]]$estimates, as.matrix(combined[[i]][,-1]))
  return(mw100_sims2[[i]])
}) 


save(mw100_sims2, file = "C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/mw100_sims2.RData")

```
 
 
```{r tidy mord1}
 broom::tidy(m100_ord1) %>% # tidy up m100_ord1 for context
  # z95 comes with stevemisc for precise values coinciding with areas under N(0,1)
  # also, these lwr/upr bounds make less sense for alphas, but we're not going to belabor those.
  mutate(lwr = estimate - p_z(.05)*std.error,
         upr = estimate + p_z(.05)*std.error) %>%
  mutate(category = "Ordinal Logistic Regression Summary") -> tidym100_ord1

simm100_ord1 %>% # we're going to name things to match up nicely with the above output
  gather(term, value, everything(), -sim) %>%
  group_by(term) %>%
  summarize(estimate = mean(value),
            `std.error` = sd(value), # you can laugh now at calling this a std.error, but I'm going somewhere with this
            lwr = quantile(value, .025),
            upr = quantile(value, .975)) %>%
  mutate(category = "Model Simulation Summary") %>%
  # this is why i'm standardizing column names...
  bind_rows(tidym100_ord1, .) %>%
  # group_by fill to drop the alphas
  group_by(term) %>%
  fill(coef.type) %>%
  filter(coef.type == "location") %>%
  ungroup() %>%
  mutate(term = fct_recode(term,
                           "Income Inequality" = "gini_2004_6_cent",
      "GDP/capita" = "gdppc_2004_6_scale",
      "Age" = "age_cent",
      "Employed" = "employed1",
      "Male" = "male1",
      "Perceived police effectiveness" = "police_effective",
      "Income quartile" = "income_quartile")) %>%
  mutate(term = fct_rev(fct_inorder(term))) %>%
  # plot..
  ggplot(.,aes(term, estimate, ymin=lwr, ymax=upr,color=category, shape=category)) +
  theme_steve_web() + post_bg() +
  scale_colour_brewer(palette = "Set1") +
  geom_hline(yintercept =  0, linetype="dashed") +
  geom_pointrange(position = position_dodge(width = .5)) + coord_flip()+
  labs(title = "Mixed Ordinal Regression on Victimization",
       x = "", y = "Coefficient (with 95% Intervals)",
       # subtitle = "Nothing here is terribly surprising and simulating allows another means to summarize uncertainty around the parameters.",
       shape = "", color="",
       caption = "Data: ICVS,PWT,SWIID. Random effect of country omitted for presentation.\nHigher values = more reported crime victimization.")
```

 

```{r, eval=FALSE}
library(brms)
# 4 chains, takes 3-4 days
m_brm <- brms::brm(ordered(num_victim_5yr) ~ age_cent + gdppc_2004_6 + (1|country), data = mod_joined1, family = 'cumulative', prior = set_prior('normal(0, 3)'))
```

```{r brms-model, eval=FALSE}
m100_brms <- brms::brm(ordered(num_victim_5yr_winz) ~ gini_2004_6_cent + gdppc_2004_6_scale + age_cent + employed + male + police_effective + income_quartile + (1 | country), family = cumulative,
  data = mod_joined1)
```



