---
title             : "The Relationship between Inequality and <br> Consumption of Security Products"
shorttitle        : "Inequality and Security Consumption"
date: "`r Sys.Date()`"

author: 
    - Dallas Novakowksi^[University of Calgary, Haskayne school of Business, Corresponding author, dallas.novakowski1@ucalgary.ca]
        #   - Conceptualization
    #   - Writing - Original Draft Preparation
    #   - Writing - Review & Editing
    #   - Analysis
    - Mehdi Mourali^[University of Calgary, Haskayne school of Business]    

# https://tinyurl.com/ybremelq for nature guideline for summary paragraph

abstract: |
  A growing body of evidence suggests that economic inequality causes humans to take more risks and engage in aggressive behaviours. In a 'winner-takes-all' environment, risky activities such as gambling, lying, and crime can be a person's only means of keeping up with competitors and accessing contested resources. There is comparatively little research investigating whether people anticipate risk-taking and conflict from their neighbors in unequal environments. An informative context for examining the distrust-inducing effects of inequality is the security market, which offers goods that protect customers from the actions of malicious agents. For instance, barred windows are purchased because they (purportedly) protect consumers from break-ins. Given the costly nature of these security products, security consumption can only be worthwhile if a consumer expects that other people have harmful intentions.

  The proposed research will examine whether inequality will increase consumers' willingness to purchase security products through two studies: 1) experimental manipulations of inequality and distributional fairness in an economic game context, and 2) multilevel analyses of the effects of nation-level inequality (k = 32) on individuals' consumption of security goods (n = 79,864).

#   
# keywords          : "inequality, security, consumption, trust, distrust, multilevel analysis, International Crime Victimization Survey"
# wordcount         : "X"

output:   
  bookdown::pdf_book
    # keep_tex: yes   # useful for troubleshooting and journal submissions 
    
#   bookdown::word_document2:   # knits to word, but doesn't work with KableExtra
#     toc: true
# always_allow_html: true

bibliography: references.bib
csl: nature.csl

latex_engine: pdflatex
linestretch: 1.5
fontsize: 12pt
indent: true
include-before:   
  - '`\newpage{}`{=latex}'  #Creates a page break in between title page and toc.

header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
-  \newcommand{\beginsupplement}{
    \setcounter{table}{0}  
    \renewcommand{\thetable}{S\arabic{table}} 
    \setcounter{figure}{0} 
    \renewcommand{\thefigure}{S\arabic{figure}}
  }
---

```{r setup, include = FALSE}
library(tidyverse)
library(kableExtra)
library(english)

#trackdown::download_file(file="docs/05_icvs_NHB_RR_stage1.Rmd")
# trackdown::update_file(here::here("docs","inequality_security_NHB_RR_stage1.Rmd"),hide_code=T)
  
load("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/iv_2005.RData") # icvs objects
load("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/icvs_solo_datafiles.RData")  # loads indidividual dataframes for different icvs configurations

load(here::here("output", "icvs_output.RData")) #icvs models, tidy tables, convergence stats
load(here::here("output", "pretest_output.RData"))    # Experiment power analyses and pretest output 
load(here::here("output", "manuscript_objects.RData"))  # design tables and hypotheses

source(here::here("scripts", "manuscript_funs.r"), local = knitr::knit_global())
# source(here::here("scripts","hypotheses.R"), local = knitr::knit_global())
```

```{r analysis-preferences, include = FALSE}
# Seed for random number generation
set.seed(1234)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = FALSE)

```

\newpage

# Introduction

<!--# 1.Inequality 2.Inequality risk-taking 3.Inequality high stakes 4. High stakes can't afford to sit out 5. (compositional - poverty) 6.(Psychosocial - relative deprivation) 7. Psychosocial relative fitness and positional bias 8. Self-protection and trust -->

<!--# 1. security 2. security and trust 3. Inequality and trust 3.1. Inequality and risk 4. Inequality and security -->

Income and wealth inequalities have risen sharply in the US in the past 30 years [@horowitz2020], this trend is of particular importance because a growing body of evidence suggests that economic inequality can lead to undesirable consequences. At a societal level, inequality has been associated with higher nation-level homicide rates [@daly2016], consumer debt [@christen2005], and reduced cooperation [@burton-chellew2013 ; @anderson2008], while at an individual level, inequality causes people to be more likely to resort to risky, hostile, and criminal activities in desperate attempts to compete for otherwise inaccessible positions and resources [@mishra2015; @payne2017; @krupp2018].

Two accounts have been proposed to explain how population-level inequality influences risk-taking in individuals. Both of which follow from risk-sensitivity theory, which predicts that organisms will take risks when more predictable strategies cannot meet the organism's goals (e.g., starvation thresholds). Accordingly, inequality fosters risk-taking by increasing the distance between individuals and their goals. First, inequality increases risk-taking through *absolute deprivation*; unequal societies tend to have more poverty, and poor people are more likely to take risks to meet their basic needs [@pridemore2008]. Second, risk-taking can be a response to envy (or *relative deprivation*), where even well-off individuals feel disadvantaged when compared to others, then take risks in an attempt to keep up with peers and a socially-defined reference point [@mishra2017].

Despite evidence for inequality's role in increasing risk-taking, simulation, cross-country, and panel data suggest that inequality also fosters a special case of risk aversion: low social trust [@knack1997; @zak2001; @berggren2006; @leigh2006; @bjørnskov2007; @wilkinson2010; @elgar2010; @barone2016; @decourson2021]. For instance, among OECD countries, inequality is associated with lower country-level trust, as computed from the World Values Survey [@barone2016]. Several theoretical streams have emerged to explain how inequality reduces trust. Each of these explanations are limited to certain contexts; for example, ranging from changes in intergroup contact, opportunity costs of vigilance, and incentives for antisocial behaviour [@jordahl2007].

A good example of how inequality impacts distrust in populations is found in a simulation study by Decourson and Nettle [@decourson2021]. They show how, in unequal environments, worse-off individuals will tend to be in poverty, and will try to exploit cooperative groups (as suggested by risk-sensitivity theory), leading others in a population to eventually avoid cooperation altogether. This theory of inequality-distrust is explicitly focused on the effects poverty (i.e. absolute deprivation). However, in many instances, poverty and inequality do not co-occur. For instance, some workplaces and neighborhoods have little to no absolute deprivation, but they can still exhibit substantial stratification; in such a circumstance, this compositional model of inequality would not predict any reduced trust. We argue that considerations of others' *relative deprivation* can help to extend this inequality-trust theory to other important contexts.

People appear to demonstrate an aversion to being the subject of social comparisons and feelings of envy. Data from small-scale preindustrial societies suggests that many individuals hold "evil eye beliefs;" superstitions that a mere envious glance can cause material harm. Moreover, these beliefs are most prevalent in societies with high wealth inequality [@gershman2015]. Likewise, a major feature of scandinavian culture is the "Jante Law" (or Janteloven/Jantelagen), which explicitly discourages individual superiority, ambition, and nonconformity. e.g., "You musn't think you are anything special." [@sandemose1936] Anglo-Saxon societies refer to the 'cutting down' of 'tall poppies', and in Asian cultures, 'the nail that stands out gets hammered down.'

If envious peers are more likely to act spitefully and take risks, being able to anticipate or detect disparities among peers may be helpful in protecting oneself from harmful behaviours. These patterns of envy aversion seem adaptive when considering the risk- and conflict- inducing effects of inequality. Indeed, theories have been proposed that predict a "fear-of-envy" equilibrium, where concerns over partner's destructive envy can reduce people's likelihood to engage in wealth-creating behaviours [@gershman2014; @brams2005].

This project will operate under the assumption that the purchase and use of security products is, at least partly, a consumer manifestation of distrust. Currently, most research on the inequality-trust relationship is based on attitudinal measures (REF). While informative tools for otherwise unobservable internal states, these psychological measures provide little information on how trust and distrust manifest in people's behaviours. Security products may improve our understanding of distrust because security products are one of the few goods that confer little intrinsic value to consumers. Whereas vacations are inherently satisfying, the costs of burglar alarms, sturdy locks, and security cameras are often only worthwhile *if consumers expect that other agents may cause them harm*. This project aims to test an envy aversion model of distrust and security consumption, which predicts that **H1)** `r hypotheses[1]` (see Table \@ref(tab:design-table) for hypotheses and design table) because **H2)** people expect disadvantaged peers to experience envy.

This theory of envy avoidance also addresses the role of fairness in shaping an individuals' experience envy. Benign envy is associated with a motivation to improve oneself, and occurs when an individual believes that the envied person's advantage is deserved. By contrast, malicious envy is characterized by a desire to harm the comparison target, and is elicited when the individual believes that their environment is unfair [@vandeven2012]. Participants are more likely to engage in helping behaviours when they believe that their partner is frustrated by the disparity and harboring antagonistic motivations (i.e., experiencing malicious envy)[@vandeven2010]. In contrast, participants who earned an advantaged position (which is associated with benign envy) tend to be less generous [@hoffman1994]. As such, this project will also test the prediction that **H3)** `r tolower(hypotheses[2])`, and that **H4)** Individuals will consume more security when inequality is the apparent result of group-based favoritism.

This theory of envy aversion makes similar directional predictions as some existing inequality-distrust theories, but makes these predictions under more generalized circumstances. Even in the absence of absolute poverty, larger material incentives for antisocial behaviour, or different opportunity costs, this model would still predict that inequality increases distrust and security consumption, out of the person's efforts to protect themselves against envious reactions. The mechanism of envy aversion is specific to predicting reduced trust towards less wealthy individuals among a reference group. This theory will be tested through a multimethod project of four experiments (Study 1) and a multilevel analysis using the International Crime Victimization Survey (ICVS; Study 2) [@vankesteren2010] to examine the relationship between economic inequality and individual consumption of products that protect users against their neighbors.

# Methods

<!-- **The Methods section in Letters and Articles should ideally not exceed 3,000 words but may be longer if necessary** -->

For all of the proposed studies, data collection and analysis will not be performed blind to the conditions of the experiments. Given that the experiments have been programmed, and analyses scripts prepared, the data collection and analysis can be completed in a week after approval; we likewise propose a resubmission date of three weeks following a successful Stage 1 review.

## Ethics information

All procedures for this project have been approved by the University of Calgary Conjoint Faculties Research Ethics Board (REB 20-1860). For the experimental studies, informed consent will be obtained for all participants, who, unless otherwise noted, will be compensated at a rate of \$12.48 CAD/hour. For multilevel analyses, the datasets are archival and publicly-available. As such, ethics approval and informed consent was obtained, and compensation disbursed, by the original researchers.

## Study 1 - Experimental studies in Inequality

### Proposed sample characteristics (Sampling plan; Study 1)

Participants will be recruited through the crowdsourcing platform Prolific Academic. Sampled participants will be filtered for fluency in English, balanced for an equal gender representation, and only eligible if they have not completed an earlier experiment in this series. After data collection, participants will be screened based on an attention check and one item testing their task comprehension. After data collection, participants will be excluded from analyses unless they pass the attention and comprehension check items. Participants will also be excluded if they have missing data for any of the measures. Erroneous or missing data will result in the participants' data excluded from analyses

<!--# Regardless of sampling method, you must list all criteria for data inclusion and/or data exclusion and how this affects your sampling strategy. This includes a full description of proposed sample characteristics. Procedures for objectively defining exclusion criteria due to technical errors or for any other reasons must be specified, including details of how and under what conditions data would be replaced. These details must be summarized in the mandatory Design table (Table @ref(tab:design-table)). -->

### Experimental procedures (Design; Study 1)

The Prolific-recruited participants will complete the studies on oTree [@chen2016], an open-source platform for conducting single- and multiplayer experiments and surveys. The experiments will be conducted in the context of two-player economic game, called the security game [@tambe2011]. In this modification of the security game, participants are paired with an anonymous partner, with each being given an endowment of funds (e.g., \$2 each). Then, participants are told that their partner will have the opportunity to try to steal money from them (e.g., \$1 at a 50% chance of success, with failed thievery attempts instead making the partner lose \$1). After learning this information, participants can spend some of their money on security products (e.g., \$0.02 per unit), with each unit of security visibly reducing the probability of a successful thievery attempt (1% reduction per unit).

The security game experiments will take place across five studies, testing the effects of inequality on security consumption (1a), with real monetary payoffs (1b). The last two experiments will assess the effects of inequality when endowments are due to one's own efforts (versus random; 1c), and whether apparent group-based discrimination increases security consumption under conditions of inequality (1d). In each study, participants will be randomly assigned to a between-subjects conditions.

Economic games examining self-protection are not new. Most notably, McEvily[@mcevily2012] devised a behavioural distrust game, whereby participants choose whether to incur a cost that guarantees that their partner splits a surplus income equally. Compared to the distrust game, the security game is explicitly probabilistic - partners have a risky choice to take funds for themselves, and security expenditures reduce the probability of successful "attacks."

### Pilot data (Study 1)

A within-subjects pretest with an initial sample of `r pretest_initial_n` crowdsourced (Prolific) participants was conducted to calibrate the price of security (\$0.02 vs. \$0.04) and the baseline likelihood of successful thefts (50%, 60%, 75%). The pilot had two attention checks and implemented a relaxed comprehension check (selecting one of two options out of five), as only 67% of participants correctly answered a single-item comprehension check, leaving n = `r screened_n` retained participants. As described in the procedure above, participants were told they would be paired with a hypothetical partner, and each would be given \$2. Then, participants were told that this partner would have the opportunity to steal \$1 from them, at a baseline chance of success, with unsuccessful thefts instead making the partner lose \$1. Then, participants were asked to indicate how much of a security product they would like to purchase, with each unit of the product costing money, and reducing the likelihood of a successful theft by 1%.

A 2w\*3w within-subjects ANOVA failed to detect significant effects of baseline likelihood of successful thefts, `r anova_extract(pretest_anova,"prob")`, nor security price, `r anova_extract(pretest_anova,"cost")`; see Figure \@ref(fig:pretest-splitplot) and Table \@ref(tab:pretest-anova)). This project will use a baseline 50% chance of successful thefts, and each unit of security costing \$0.02. This condition also seems to produce the least skewed data, at `r round(base_skew,2)` (e.g., compared to skew = `r round(skew_groups["75_02"],2)` when probability is 75% and price \$0.02).

<!-- participants has already been conducted to find which values of base successful thefts and security prices yielded a favourable distribution in security consumption, while avoiding ceiling/floor effects. After examining the data (Figure @ref(fig:pretest-splitplot)), baseline probabilities of successful thefts and security price had  -->

### Measurement and variables (Study 1)

Throughout these experiments, participants will be asked to choose how many units of a security product they wish to consume, ranging from 0 to 50. For the purposes of analysis and interpretation, security spending will serve as the final dependent variable where possible. Security spending will be computed by multiplying the number of security units consumed by the good's per-unit cost. Before and after participants make their security decision, they will be asked a battery of self-report questions about the game. For example, before their security decision, participants will be rate their agreement with the statements, "my partner is probably going to try stealing from me", and "my partner probably feels jealous of me."

Given the constraints on the minimum and maximum values for the variables in the experimental designs, outliers are considered legitimate values and will not be deleted. Instead, influential observations will be winzorized to the 5th and 95th percentile values (using r function `DescTools::Winsorize`). Influential observations will be flagged as having cook's distance values exceeding the threshold 4/n; where n = \# of observations.

### Study 1a: Inequality and security spending

The first experiment will use a mixed (2b\*2w) design, manipulating high versus low economic inequality between subjects in a hypothetical context, to test whether **H1)** `r tolower(hypotheses[1])` (Table \@ref(tab:design-table)), versus **H1~0~)** no significant effect of inequality. Participants will be asked to imagine that they are paired with another participant, where they will each be assigned to different roles (Blue or Yellow), and endowed with money. In the equal condition, participants will be told that their partner received the same amount of funds as themselves (\$2). In the unequal condition, participants will be told that their partner instead received only \$1. To improve power and test the generalizability of the effect of inequality, stake size will be manipulated within-subjects (e.g., \$2 vs. \$200). As such, participants will provide two sets of responses for game-related self-report items and security consumption decisions.

Before participants make their security decisions, they will be asked four questions on whether they think their partner would be envious of them (e.g., "My partner probably feels envious of me"). These items will be summed and will be used to test whether **H2)** `r hyp2_min[1]` versus **H2~0~)** no significant indirect effect of anticipated partner envy from inequality to security consumption. This procedure has a notable strength by controlling the participant's income across levels of inequality (as the partner's money is the only income that changes). Results will be analyzed using a mixed model and mediation analyses (`afex::aov_ez`, `mediate::mediate`).

#### A priori power analysis (Study 1a)

Using mean and standard deviation values of security expenditure (*M*=`r round(base_mean,2)`, *SD*=`r round(base_sd,2)`) from the pretest, an a priori power analysis has been conducted using r package SuperPower [@lakens2021a] (see Figures \@ref(fig:mixed-design) and \@ref(fig:powerplot2) and and Tables \@ref(tab:2b2w-anova-exact) and \@ref(tab:2b2w-pairwise-power)). Given the unknown size of the hypothesized effect, this study will target an effect size of cohen's f = `r round(security_exact_5$main_results$cohen_f[1],2)`, which corresponds to a pattern of means that yields a cohen's d = .2 in an independent-samples t-test; a small effect according to statistical convention [@cohen1988]. To achieve 95% power with one tail, studies must have a minimum of n = `r round(cell_n_5,0)` participants per cell, or a study-wide total of n = `r format(round(cell_n_5*2,0),big.mark = ",", scientific = FALSE)`.


### Study 1b: Inequality with real payoffs

The use of hypothetical scenarios in Study 1a may limit the reliability of any results. For instance, compared to hypothetical rewards, participants take fewer risks when incentivized by real payoffs and experiencing negative feedback [@xu2016; @hertwig2001]. Study 1b will serve to replicate Study 1a, with the sole difference being that **H1)** and **H1a)** are being tested when participants believe they are playing with a real partner, and that their decision could have real payoffs. Specifically, participants will be told that they will be entered in a lottery where three participants will be selected to receive the payoff from one of their decisions.

This study will also use a mixed (2b\*2w) experimental design, manipulating inequality between subjects, and the starting probabilities of successful thefts (50% vs. 60%) within subjects. Study 1b will determine its targeted sample size by seeking 95% power for the observed main effect of inequality in Study 1a. If Study 1a fails to find a significant effect of inequality, Study 1b will target an effect size of cohen's f = `r round(security_exact_5$main_results$cohen_f[1],2)` (n=`r format(round(cell_n_5*2,0),big.mark = ",", scientific = FALSE)` . Results will be analyzed using a mixed ANOVA and mediation analyses. In order to test **H1)**'s robustness to stake size, an Equivalence Hypothesis Test will be conducted on the effect of stake size, using the 'two-one-sided t-test' (TOST) approach (Schuirmann, 1987).



### Study 1c: Meritocracy

Groups will always be characterized by inequality through some type of currency, whether wealth, income, or status. Indeed, a popular idea is that inequality propels individuals towards productive behaviours (REF). In this line of reasoning, the presence of inequality can be less of an issue compared to how the inequality was produced. The next experiment will test whether **H3)** `r tolower(hypotheses[2])` (Table \@ref(tab:design-table); versus **H3~0~)** no significant inequality\*merit interaction). As planned comparisons, Study 1c will also examine whether **H3a)** hi-inequality/random income condition will have significantly higher rates of security consumption than all other conditions (i.e., **.1)** hi-inequality/merit income, **.2)** no-inequality/random income, & **.3)** no-inequality/merit income. Study 1c again varies inequality (yes/no) as the previous studies, but also manipulates whether one's income is seemingly allocated *randomly or based on task performance*.

For this random vs. merit manipulation, all participants will complete a slider task, where they have one minute to correctly place 24 sliders at the midpoint of various lines that appear on their screens. Participants will then either be told that their incomes were allocated randomly, or because they correctly placed more sliders than their partner. The slider task has been used in several studies to examine the effects of effort in distributional preferences, and has demonstrated that individuals that win their  income through performance are less likely to support redistribution of effort-driven income allocations [@cassar2019].

As a manipulation check, before making their consumption decision, participants will be asked to report whether they believed the incomes were disbursed fairly (e.g.,"The way incomes were given for this game is fair"). Participants will be excluded from analyses if they complete fewer than 3 sliders, which would both indicate low engagement in the study and likely undermine the credibility of the manipulation. Results will be analyzed using a 2b\*2b between-subjects ANOVA.


<!--# competing hypothesis: deserved superiority will lead to a greater desire to retain relative positioning –> more consumption - could nullify or reverse -->

#### A priori power analysis (Study 1c)

Study 1c requires sufficient statistical power to detect a "no-way" interaction. Given the unknown size of the hypothesized inequality\*merit interaction, this study will target an effect size of cohen's f = .1. To achieve 95% power, studies must have a minimum of `r cell_n_2b2b_f01` participants per cell, or a study-wide total of n=`r format(cell_n_2b2b_f01*4,big.mark = ",", scientific = FALSE)`. This large sample size is necessary given the possibility of a "no-way" interaction, where only one condition is significantly different from the remaining conditions [@simonsohn2015].

In this case, the simple effect of inequality is nullified depending on the moderator, and thus requires at least twice as many subjects per cell to achieve as much power as the simple effect. Figures \@ref(fig:2b2b-design) and \@ref(fig:2b2b-powerplot) display the pattern of means and per-condition sample size thresholds to obtain 95% power. Additionally, see Tables \@ref(tab:2b2b-anova-power) and \@ref(tab:2b2b-pairwise-power) for the power and effect sizes of the ANOVA and pairwise comparisons, respectively.

### Study 1d: Group membership

As described in the literature review to Study 1c, fairness appears to have its roots in avoiding partiality. Since Study 1c directly compares meritocratic versus random allocation of resources, it does not answer the more fundamental question: do people consume more security goods when inequality appears to be more discriminatory and unfair? Study 1d will keep a constant presence of inequality, but manipulate whether the partner is a member of an ingroup or outgroup, and whether the incomes are distributed randomly, or by a separate *ingroup individual*. Study 1d will use a 2b\*2b between subjects ANOVA to test whether **H4)** `r tolower(hypotheses[3])` (versus **H4~0~)** no significant outgroup\*agentic-distribution interaction). Participants will complete the same fairness manipulation check as and be powered for the same effect as in Study 1c.

## Study 2 - Multilevel Analysis of the ICVS

The highly-controlled nature of the experiments in Study 1 allow for precise tests of the hypothesized inequality-security consumption relationship. Although they offer real consequences for participants' choices, economic games are highly artificial by design. Insights generated in the security game may not be generalizable to consumers' daily consumption context. To provide a test with greater external validity, this project supplements experimental results with the multilevel analysis of archival data assessing respondents' use of security products, to test whether **H5)** `r tolower(hypotheses[4])` (versus **H5~0~)** no significant effect of nation-level inequality).

The format of a registered report carries substantial benefit when analyzing archival data, as such data-dependent analysis, the so-called "garden of forking paths"[@gelman2014] allows a researcher to pick and choose those data and analyses that best suit their preferred conclusions[@simmons2011]. The commitment to reprocessing and analysis plans before model estimation helps to reduce these researcher degrees of freedom.

### Study 2 - experimental procedures (Design)

The relationship between country-level inequality and the consumption of security products will be tested using a multilevel ordinal regression[@christensen2019]. This analysis will be accomplished using a combination of three archival datasets: the ICVS[@vankesteren2010], the Standardized World Income Inequality Database (SWIID)[@solt2020], and the Penn World Tables (version 10.0; PWT)[@feenstra2015].

<!-- Atkinson and Brandolini (2001) discuss these and several related pitfalls in their excellent survey on the use of secondary data sets in studies of income inequality. -->

### Study 2 - Proposed sample characteristics (Sampling plan)

First, indicators of security consumption have been accessed from the ICVS [@vankesteren2010], which has surveyed over 300,000 people across `r cleaned_countries` different countries on householders' experiences with crime, policing, crime prevention, and feelings of unsafety. Second, nation-level inequality in disposable income was accessed through the SWIID[@solt2020]. The SWIID has been designed to maximize the comparability of income inequality data while maintaining the widest possible coverage across countries and over time. Lastly, countries' expenditure-side real GDP and populations were retrieved from the PWT[@feenstra2015].

Although the ICVS, PWT, and SWIID extend across multiple years, the current analysis will be limited to observations across 2004-2006. After excluding for missing data on security and key independent variables, k = `r sum(summary(iv_2005$country)>0)` countries, and n = `r format(nrow(iv_2005), big.mark = ",", scientific = FALSE)` participants were retained for primary analyses. Further screening information can be found in the supplementary materials, but as a summary, this sample specification provides a balance of relative data recency, larger sample size, and a larger range of values on key variables such as inequality and victimization experiences.

<!-- comment on confounds between security measures and home ownership -->

### Study 2 - Measurement and Variables

#### Outcome variable

```{r prev-var-recode, echo = FALSE}
prevention_min2 <- dplyr::recode(prevention_min2, "prev_burglar_alarm" = "burglar alarm", "prev_special_door_locks" = "special door locks", "prev_special_grills" = "window grills", "prev_high_fence" = "high perimeter fence", "prev_caretaker_security" = "presence of a caretaker who provides security")

```

Security consumption (*M* = `r round(mean(iv_2005$total_security),2)`, *SD* = `r round(sd(iv_2005$total_security),2)`) has been computed by summing respondents' self-reported ownership of `r as.english(length(prevention_min2))` different preventative measures: `r glue::glue_collapse(x = prevention_min2, ", ", last = ", and ")`. This variable excludes several security items in the ICVS due to missing data, such as owning a watch dog, having surveillance arrangements with neighbors, and purchasing insurance against criminal activities.

<!-- a burglar alarm, special door locks, special grills, a high perimeter fence, and caretaker security. -->

#### Predictor variables

The focal model will include predictor variables at country- and individual-levels: at the nation level, 1) nation-level inequality (gini - disposable income), 2) expenditure-side GDP per capita (matched purchasing power parity), and 3) nation-level self-reported crime victimization; and at the respondent level: 1) age (five-year increments), 2) gender (`r round(summary(iv_2005$gender)["female"]/nrow(iv_2005)*100,0)`% female), 3) employed (`r round(summary(iv_2005$employed)[1]/nrow(iv_2005)*100,0)`% employed), and 4) number of individual victimization experiences. See Table \@ref(tab:s2desc-tab) for remaining descriptive statistics.

As an index of inequality, the Gini coefficient is most intuitively defined as the average difference in income between all pairs in a population, divided by twice the average income in the population. The SWIID consists of 100 separate dataframes, with varying gini values to reflect the values' uncertainty (i.e., larger standard errors) when making comparisons across different countries or time periods. The PWT-selected GDP values are adjusted for Purchasing Power Parity, "to compare relative living standards across countries at a single point in time" [@feenstra2015]. Nations' GDP values were divided by their population sizes to yield a per-capita GDP value, which will be used for these analyses.

```{r vic-var-recode, echo = FALSE}
vic_var_assault <- dplyr::recode(vic_var_assault, "cartheft_5yrs" = "car thefts", "motortheft_5yrs" = "motor vehicle thefts", "bicyctheft_5yrs" = "bicycle thefts", "burglar_5yrs" = "burglaries", "attempt_5yrs" = "attempted burglaries", "robbery_5yrs" = "robberies", "pers_theft_5yrs" = "thefts", "assault_5yrs"="physical assaults")
```

Self-reported crime victimization will likewise be computed as a count variable for the following items: `r glue::glue_collapse(x = vic_var_assault, ", ", last = ", and ")`, summing `r as.english(length(vic_var_assault))` binary items on respondent's experiences of victimization. This variable will be represented twice in the analyses - once as a nation-average value, and again for the individual respondent.

### Proposed analysis pipeline, including all preprocessing steps, precise description of all planned analyses

Since the ICVS, SWIID, and PWT are publicly available, preprocessing and inclusion/exclusion steps have already been implemented, and are available in the sourced and attached code. A notable preprocessing steps is that gini and GDP values were averaged across 2004-2006 into a single index, to reflect the years where the survey sweeps were conducted.

Analyses will predict security consumption as an ordinal variable (0: no security consumption, 1: one unit of security consumption, 2: two units, or 3: three or more units consumed), using a cumulative link mixed model. The outcome variable exhibits excessive skewness, and is not suitable to a linear model. Likewise, despite security consumption being a count variable of multiple binary occurrences, it is unlikely to meet the assumptions for a poisson process, which assumes that the maximum value stretches out indefinitely.

The coefficient estimates and standard errors will be extracted through simulations. The distributions for fixed-effects estimates will be generated by simulation from 100 dataframes, which each consist of coefficients and variance/covariance matrices from a model. These distributions will be used to calculate point estimates and standard errors for each fixed effect. If the weighted ordinal regression fails to converge, the model will be conducted without frequency weights. Likewise, if the unweighted ordinal regression fails to converge, a weighted mixed logistic regresssion will be conducted. If the weighted mixed logistic regression fails to converge, an unweighted mixed logistic regression will be conducted instead. In addition, supplementary analyses will assess whether the effects of inequality persist in a smaller, variable-maximizing dataset (k = `r sum(summary(iv_2005_mod$country)>0)`, n = `r format(nrow(iv_2005_mod), big.mark = ",", scientific = FALSE)`).

### Pilot data (Study 2)

As a proof of concept, the proposed analyses have been conducted on a different variable: individuals' victimization experiences. Like security behaviours, victimization is operationalized by adding together multiple yes/no items, resulting in a similar distribution to security behaviours (see Figures \@ref(fig:icvs-security-hist) and \@ref(fig:icvs-victim-hist)). In this case, `r model_nonconvergence` of 100 analyzed dataframes failed to converge (preventing calculation of standard errors). As such, mixed ordinal regression without frequency weighting was conducted, and had no convergence issues.

In this ordinal mixed regression model, gini has a significant relationship with personal experiences of victimization, `r tidy_extract(iv_tidy_ci,"gini_2004_6_cent")`, corresponding to a `r iv_tidy_ci[3,"estimate"]` increase the natural logged odds of observing a 2 versus a 1 or 0 (and 1 versus 0), per unit change in the gini coefficient. See Figure \@ref(fig:icvs-victim-dotwhisk-iv) and Table \@ref(tab:s2-victim-table-iv) for more results. In addition, see Figures \@ref(fig:icvs-victim-dotwhisk) and \@ref(fig:icvs-victim-dotwhisk-min), as well as Tables \@ref(tab:s2-victim-table) and \@ref(tab:s2-victim-table-min) for supplementary analyses on the effects of a larger model (i.e., including assaults, income quartile, and perceived police effectiveness).

<!--# Full descriptions must be provided of any outcome-neutral criteria that must be met for successful testing of the stated hypotheses.Such quality checks might include the absence of floor or ceiling effects in data distributions, positive controls, or other quality checks that are orthogonal to the experimental hypotheses. -->

<!-- https://marissabarlaz.github.io/portfolio/ols/ -->

<!-- e expect a -1.095 increase (or a 1.095 decrease) in the expected value of rating on the log odds scale, given all of the other variables in the model are held constant. In other words, when going from object to subject, the likelihood of a 4 versus a 1-3 on the rating scale decreases by 1.095 on the log odds scale, the likelihood of a 3 versus a 1-2 on the rating scale decreases by 1.095 on the log odds scale, and the likelihood of a 2 versus a 1 on the rating scale decreases by 1.095 on the log odds scale. -->

<!-- If we want to look at this on the odds scale, we can exponentiate using exp() to look at odds. -->

\newpage

# Data availability statement

Raw data can be accessed at [\<https://github.com/DallasNovakowski/inequality_security\>](https://github.com/DallasNovakowski/security_game){.uri}. Archival sources available at the following links: [Penn World Tables (PWT; for GDP)](https://www.rug.nl/ggdc/productivity/pwt/?lang=en), [The Standardized World Income Inequality Database (SWIID - for inequality)](https://fsolt.org/swiid/), and [International Crime Victims Survey (ICVS - for crime and security measures)](https://easy.dans.knaw.nl/ui/datasets/id/easy-dataset:44578).

# Code availability statement

Interactive demos of experimental games can be accessed at [(https://security-pretest.herokuapp.com/demo](https://security-pretest.herokuapp.com/demo){.uri}), and experimental code can be accessed at [\<https://github.com/DallasNovakowski/security_game\>](https://github.com/DallasNovakowski/security_game){.uri}. Preprocessing and analysis code is available at \<<https://github.com/DallasNovakowski/inequality_security>\>.

\newpage

# References

\begingroup

::: {#refs custom-style="Bibliography"}
:::

\endgroup

\newpage

# Acknowledgements

The funders had no role in study design, data collection and analysis, decision to publish or preparation of the manuscript.

# Author contributions

All authors edited the manuscript and approved the final version. D.N. designed the studies and wrote the manuscript, in consultation with M.M.. D.N. collected, processed, and analysed the data. The study designs were based on theory developed by D.N.

# Competing interests

The authors declare no competing interests.

\newpage

<!-- # Figures & Figure captions -->

<!-- \newpage -->

# Tables

```{r design-table}
kd <- knitr::kable(design, caption = 'Design table', format = "latex",
                   # booktabs = TRUE, 
                   col.names = design_columns) %>%
  kable_styling(full_width = FALSE) %>%
  column_spec(1,width = "1.7in") %>%
  column_spec(2,width = "1.3in") %>%
  column_spec(3,width = "1.7in") %>%
  column_spec(4,width = "1.3in") %>%
  column_spec(5,width = "1.7in") %>%
  kableExtra::landscape()

kd
```

```{r s2desc-tab}
s2desc_columns <- c("Variable",	"Mean",	"SD")

Variable <- c("gini","GDP","nation-level victimization","age",
              # "income quartile",
              "individual victimizations",
              # "police effectiveness",
              "security consumed")
Mean <-
  c(round(mean(gd$gini_2004_6),2),
    format(round(mean(gd$gdppc_2004_6)), big.mark = ",", scientific = FALSE),
    round(mean(gd$num_victim_5yr),2),
    round(mean(iv_2005$age_num),2),
    # round(mean(iv_2005$income_quartile),2),
    round(mean(iv_2005$num_victim_5yr),2),
    # round(mean(iv_2005$police_effective),2),
    round(mean(iv_2005$total_security),2)
    )
SD <-
  c(round(sd(gd$gini_2004_6),2),
    format(round(sd(gd$gdppc_2004_6)), big.mark = ",", scientific = FALSE),
    round(sd(gd$num_victim_5yr),2),
    round(sd(iv_2005$age_num),2),
    # round(sd(iv_2005$income_quartile),2),
    round(sd(iv_2005$num_victim_5yr),2),
    # round(sd(iv_2005$police_effective),2),
    round(sd(iv_2005$total_security),2)
    )

s2_desc <- data.frame(Variable, Mean, SD)

s2_desc_tab <- knitr::kable(s2_desc, caption = 'Study 2 Descriptive Statistics',
                   # booktabs = TRUE, 
                   col.names = s2desc_columns, booktabs = T) %>%
  kable_styling(full_width = FALSE) %>%
  column_spec(1,width = "2in") %>%
  column_spec(2,width = ".6in") %>%
  column_spec(3,width = ".6in")

s2_desc_tab
```

<!-- \clearpage -->

<!-- # (APPENDIX) Appendix {-}  -->

<!-- # Appendix A -->

\clearpage

# Supplementary material {.unnumbered}

\beginsupplement

```{r, child=c('supp_materials.Rmd')}

```
