---
title             : "The Relationship between Inequality, Distrust, and the Consumption of Security Products"
shorttitle        : "Inequality and Security Consumption"
date: "`r Sys.Date()`"

author: 
    - Dallas Novakowksi^[University of Calgary, Haskayne school of Business, Corresponding author, dallas.novakowski1@ucalgary.ca]
        #   - Conceptualization
    #   - Writing - Original Draft Preparation
    #   - Writing - Review & Editing
    #   - Analysis
    - Mehdi Mourali^[University of Calgary, Haskayne school of Business]    

# https://tinyurl.com/ybremelq for nature guideline for summary paragraph

abstract: |
  A growing body of evidence suggests that economic inequality causes humans to take more risks and engage in aggressive behaviours. In a 'winner-takes-all' environment, risky activities such as gambling, lying, and crime can be a person's only means of keeping up with competitors and accessing contested resources. There is comparatively little research investigating whether people anticipate risk-taking and conflict from their neighbors in unequal environments. An informative context for examining the distrust-inducing effects of inequality is the security market, which offers goods that protect customers from the actions of malicious agents. For instance, barred windows are purchased because they (supposedly) protect consumers from break-ins. Given the costly nature of these security products, security consumption can only be worthwhile if a consumer expects that other people have harmful intentions.The proposed research will examine how inequality affects consumers' willingness to purchase security products experimental manipulations of inequality and distributional fairness in an economic game context.

#   
# keywords          : "inequality, security, consumption, trust, distrust, multilevel analysis, International Crime Victimization Survey"
# wordcount         : "X"

output:   
  bookdown::pdf_book
    # keep_tex: yes   # useful for troubleshooting and journal submissions 
    
#   bookdown::word_document2:   # knits to word, but doesn't work with KableExtra
#     toc: true
# always_allow_html: true

bibliography: ["references.bib","r-references.bib"]
csl: nature.csl

latex_engine: pdflatex
linestretch: 1.5
fontsize: 12pt
indent: true
include-before:   
  - '`\newpage{}`{=latex}'  #Creates a page break in between title page and toc.

header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
-  \newcommand{\beginsupplement}{
    \setcounter{table}{0}  
    \renewcommand{\thetable}{S\arabic{table}} 
    \setcounter{figure}{0} 
    \renewcommand{\thefigure}{S\arabic{figure}}
  }
- \usepackage{placeins}
---

```{r setup, include = FALSE}
library(tidyverse)
library(kableExtra)
library(english)

#trackdown::download_file(file="docs/05_icvs_NHB_RR_stage1.Rmd")
# trackdown::update_file(here::here("docs","inequality_security_NHB_RR_stage1.Rmd"),hide_code=T)
  
# load("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/iv_2005.RData") # icvs objects
# load("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/icvs_solo_datafiles.RData")  # loads indidividual dataframes for different icvs configurations

# load(here::here("output", "icvs_output.RData")) #icvs models, tidy tables, convergence stats
load(here::here("output", "pretest_output.RData"))    # Experiment power analyses and pretest output 
load(here::here("output", "manuscript_objects.RData"))  # design tables and hypotheses
# load(here::here("output","m100_log_iv_sims.RData"))
source(here::here("scripts", "manuscript_funs.r"), local = knitr::knit_global())
# source(here::here("scripts","hypotheses.R"), local = knitr::knit_global())
```

```{r analysis-preferences, include = FALSE}
# Seed for random number generation
set.seed(1234)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, echo = FALSE, warning=F)

```

\newpage

# Introduction

<!--# 1.Inequality 2.Inequality risk-taking 3.Inequality high stakes 4. High stakes can't afford to sit out 5. (compositional - poverty) 6.(Psychosocial - relative deprivation) 7. Psychosocial relative fitness and positional bias 8. Self-protection and trust -->

<!--# 1. security 2. security and trust 3. Inequality and trust 3.1. Inequality and risk 4. Inequality and security -->

Income and wealth inequalities have risen sharply in the US in the past 30 years,[@horowitz2020] these growing disparities are particularly important because a growing body of evidence suggests that economic inequality can lead to undesirable consequences. At a societal level, inequality has been associated with higher nation-level homicide rates,[@daly2016] consumer debt,[@christen2005] and reduced cooperation,[@burton-chellew2013 ; @anderson2008] while at an individual level, inequality causes people to resort to more risky, hostile, and criminal activities in desperate attempts to compete for otherwise inaccessible positions and resources.[@mishra2015; @payne2017; @krupp2018]

Two accounts have been proposed to explain how population-level inequality influences risk-taking in individuals. Across both explanations, inequality fosters risk-taking by increasing the distance between individuals and their goals. First, inequality increases risk-taking through *absolute deprivation*; unequal societies tend to have more poverty, and poor people are more likely to take risks to meet their basic needs.[@pridemore2008] Second, risk-taking can be a response to envy (or *relative deprivation*), where even well-off individuals feel disadvantaged when compared to others, then take risks in an attempt to keep up with peers and a socially-defined reference point.[@mishra2017] The risk-inducing effects of both absolute and relative deprivation follow from risk-sensitivity theory, which predicts that organisms will take risks when more predictable strategies cannot meet the organism's goals (e.g., starvation thresholds, and socially-influenced goal states).[@caraco1980]

Despite evidence for inequality's role in increasing risk-taking, data from simulation, cross-country, and panel studies suggest that inequality also fosters a special case of risk aversion: low social trust.[@knack1997; @zak2001; @berggren2006; @leigh2006; @bjørnskov2007; @wilkinson2010; @elgar2010; @barone2016; @decourson2021] For instance, among OECD countries, inequality is associated with lower country-level trust, as computed from the World Values Survey.[@barone2016] Trust/distrust can be considered as the polar opposites on the same psychological scale, generally reflecting our belief whether another person's actions will be beneficial versus harmful to us.[@kramer1999; @schul2008] Given that trust can impact the politicians we vote for, the products we buy, and the relationships we keep, the processes and predictors of trust has understandably attracted a lot of attention. Several theoretical streams have emerged to explain how inequality reduces trust, such as changes in intergroup contact, or larger incentives for antisocial behaviour.[@jordahl2007]

A good example of how inequality may impact distrust in populations is found in a simulation study by Decourson and Nettle.[@decourson2021] They show how, in unequal environments, worse-off individuals will tend to be in poverty, and will try to exploit cooperative groups to surpass desperation thresholds (e.g., affording rent; as suggested by risk-sensitivity theory). After observing those exploitative behaviours, others in a population become distrustful, eventually avoid cooperation altogether.

Decourson and Nettle's theory of inequality-distrust focuses on the cascading effects of poverty (i.e. absolute deprivation). However, the authors also suggest that an individual's threshold for desperation is not a fixed dollar value. Instead, a belief whether one's resources are sufficient can be shaped by comparisons to normal standards of living, in addition to comparisons with nearby social partners.[@mishra2017] This project seeks to elaborate on these considerations of *relative deprivation* (i.e., envy). In particular, our proposed extension is that inequality will lead to reduced trust without individuals needing to observe the risky or criminal activities in the population, because people will try to anticipate and protect themselves against envious reactions.

People seem to be wary of being envied. Data from small-scale preindustrial societies suggests that many individuals hold "evil eye beliefs," superstitions that a mere envious glance can cause real, material harm. These beliefs are most prevalent in societies with high wealth inequality, and especially where one's resources are easily visible and vulnerable to third parties.[@gershman2015] These patterns of envy avoidance seem adaptive when considering the risk- and conflict- inducing effects of inequality. If envious peers are more likely to be hostile and take risks, it may be helpful to anticipate when one's neighbors will feel relatively deprived, and to protect oneself from spiteful or desperate behaviours.[@gershman2014; @brams2005]

This project seeks to test an envy-anticipation theory of distrust, which generally predicts *that inequality reduces trust towards less well-off individuals among a reference group*. The hypothesized mechanism for this project is that individuals expect envy from disadvantaged partners, and that these individuals consequently make efforts to protect themselves against the possibility of peers' hostile or risky behaviours. This envy-anticipation theory of distrust makes similar directional predictions as some existing inequality-distrust theories, but makes these predictions under more generalized circumstances. Even in the absence of absolute poverty, changes in intergroup contact, larger material incentives for antisocial behaviour, or different opportunity costs, this model would still predict that inequality increases distrust and security consumption, out of the person's efforts to protect themselves against envious reactions.

This envy-anticipation theory of distrust also considers the role of fairness in shaping individuals' experiences of envy, which can vary in their underlying attributions and motivations. For instance, benign envy is associated with a motivation to improve oneself, and occurs when the person believes that the envied person's advantage is deserved. By contrast, malicious envy is characterized by a desire to harm the comparison target, and is elicited when the individual believes that their environment is unfair.[@vandeven2012] When participants believe that their partner is experiencing malicious envy, they are more likely to engage in helping behaviours.[@vandeven2010] In contrast, participants who earned an advantaged position (which is associated with benign envy) tend to be less generous.[@hoffman1994] Together, the existing research suggests that people are not just sensitive to inequality and fairness, but that their decision-making also considers and anticipates others' psychological reactions to disparities.

This project will operate under the assumption that the purchase and use of security products is, in part, a consumer manifestation of distrust. Currently, most research on the inequality-trust relationship is based on attitudinal measures.[@jordahl2007] While informative tools for otherwise unobservable internal states, these psychological measures provide limited information on how trust and distrust manifest in people's behaviours. Security products may improve our understanding of distrust because they are one of the few goods that confer little intrinsic value to consumers. Whereas vacations are inherently satisfying, the costs of burglar alarms, sturdy locks, and security cameras are often only worthwhile *if consumers expect that other agents may cause them harm*.

This project aims to test an envy aversion model of distrust and its ability to predict security consumption. Specifically, this project will test the prediction that **H1)** `r tolower(hypotheses[1])` (see Table \@ref(tab:design-table) for hypotheses and design table) because **H2)** people expect disadvantaged peers to experience envy. Addressing the moderating effects of fairness on the effects of inequality, this project will also test the prediction that **H3)** `r tolower(hypotheses[2])`, and that **H4)** Individuals will consume more security when inequality is the apparent result of group-based favoritism.

# Methods

<!-- **The Methods section in Letters and Articles should ideally not exceed 3,000 words but may be longer if necessary** -->

For all of the proposed studies, data collection and analysis will not be performed blind to the conditions of the experiments. Given that the experiments have been programmed, and analyses scripts prepared, the data collection and analysis can be completed within a week after approval; we likewise propose a resubmission date of three weeks following a successful Stage 1 review.

## Ethics information

All procedures for this project have been approved by the University of Calgary Conjoint Faculties Research Ethics Board (REB 20-1860). For the experimental studies, informed consent will be obtained for all participants, who, unless otherwise noted, will be compensated at a rate of \$12.48 CAD/hour.

## Study 1 - Experimental studies in Inequality

### Proposed sample characteristics (Sampling plan; Study 1)

Participants will be recruited through the crowdsourcing platform Prolific Academic. Sampled participants will be filtered for fluency in English, participation through computer (vs. smartphone/tablet), only eligible if they have not completed an earlier experiment in this series, and balanced for an equal gender representation. After data collection, participants will be screened based on an attention check and one item testing their task comprehension. After data collection, participants will be excluded from analyses unless they pass the attention and comprehension check items. Participants will also be excluded if they have erroneous or missing data for any measures used in analyses.

<!--# Regardless of sampling method, you must list all criteria for data inclusion and/or data exclusion and how this affects your sampling strategy. This includes a full description of proposed sample characteristics. Procedures for objectively defining exclusion criteria due to technical errors or for any other reasons must be specified, including details of how and under what conditions data would be replaced. These details must be summarized in the mandatory Design table (Table @ref(tab:design-table)). -->

### Experimental procedures (Design; Study 1)

The Prolific-recruited participants will complete the studies on oTree,[@chen2016] an open-source platform for conducting single- and multi-player experiments and surveys. The experiments will be conducted in the context of a two-player economic game, called the security game.[@tambe2011] In this modification of the security game, participants are (supposedly) paired with an anonymous partner, with each being given an endowment of funds (e.g., \$2 each). Then, participants are told that their partner will have the opportunity to try to steal money from them (e.g., \$1 at a 50% chance of success, with failed thievery attempts instead making the partner lose \$1). After learning this information, participants can spend some of their money on security products (e.g., \$0.02 per unit), with each unit of security visibly reducing the probability of a successful thievery attempt (1% reduction per unit).

The security game experiments will take place across four studies, testing the effects of inequality on security consumption (Study 1a), with real monetary payoffs (1b). The last two experiments will assess the effects of inequality when endowments are due to one's own efforts (versus random; 1c), and whether apparent group-based discrimination increases security consumption under constant of inequality (1d). Except for Study 1d, for each study, inequality will be randomly assigned as a between-subjects condition.

Economic games examining self-protection are not new. Most notably, McEvily[@mcevily2012] devised a behavioural distrust game, and found that participants are very willing to pay money to guarantee that their partner splits a surplus income equally. Compared to the distrust game, the security game is explicitly probabilistic; partners have a risky choice to try stealing funds for themselves, and security expenditures reduce the probability of successful "attacks." The risky nature of the security game is particularly useful because our theory of envy aversion focuses on the anticipation of increased risk-taking.

### Pilot data (Study 1)

A within-subjects pretest with an initial sample of `r pretest_initial_n` crowdsourced (Prolific) participants was conducted to calibrate the price of security (\$0.02 vs. \$0.04) and the baseline likelihood of successful thefts (50%, 60%, 75%). The pretest had two attention checks a comprehension check, leaving n = `r screened_n` retained participants. As described in the procedure above, participants were told they would be paired with a hypothetical partner, and each would be given \$2. Then, participants were told that this partner would have the opportunity to steal \$1 from them, at a baseline chance of success, with unsuccessful thefts instead making the partner lose \$1. Then, participants were asked to indicate how much of a security product they would like to purchase, with each unit of the product costing money, and reducing the likelihood of a successful theft by 1%.

A 2w\*3w within-subjects ANOVA failed to detect significant effects of baseline likelihood of successful thefts, `r anova_extract(pretest_anova,"prob")`, nor security price, `r anova_extract(pretest_anova,"cost")`; see Figure \@ref(fig:pretest-splitplot) and Table \@ref(tab:pretest-anova)). This project will use a baseline 50% chance of successful thefts, and each unit of security costing \$0.02 (or proportional equivalent). This condition seems to produce the lowest means and the least skewed data, at `r round(base_skew,2)` (vs. skew = `r round(skew_groups["75_02"],2)` when probability is 75% and price \$0.02).

<!-- participants has already been conducted to find which values of base successful thefts and security prices yielded a favourable distribution in security consumption, while avoiding ceiling/floor effects. After examining the data (Figure @ref(fig:pretest-splitplot)), baseline probabilities of successful thefts and security price had  -->

### Measurement and variables (Study 1)

Throughout these experiments, participants will be asked to choose how many units of a security product they wish to consume, ranging from 0 to 50. For the purposes of analysis and interpretation, security spending will serve as the final dependent variable where possible. Security spending will be computed by multiplying the number of security units consumed by the good's per-unit cost. Before participants make their security decision, they will be asked a battery of self-report questions about the game. For example, before their security decision, participants will rate their agreement with the statements, "my partner is probably going to try stealing from me", and "my partner probably feels jealous of me."

Given the constraints on the minimum and maximum values for the variables in the experimental designs, outliers are considered legitimate values and will not be deleted. Instead, influential observations will be winzorized to the 5th and 95th percentile values. Influential observations will be flagged as having cook's distance values exceeding the threshold 4/n; where n = \# of observations.

### Study 1a: Inequality and security spending

The first experiment will use a mixed (2b\*2w) design in a hypothetical context, manipulating economic inequality (yes vs. no) between-subjects, and stake size (\$2 vs. \$300) within-subjects, to test whether **H1)** `r tolower(hypotheses[1])` (Table \@ref(tab:design-table)), versus **H1~0~)** no significant effect of inequality. Participants will be asked to imagine that they are paired with another participant, where they will each be assigned to different roles (Blue or Yellow), and endowed with money. In the equal condition, participants will be told that their partner received the same amount of funds as themselves (e.g., \$2). In the unequal condition, participants will be told that their partner instead received only half their endowment (e.g. \$1).

Before participants make their security decisions, they will be asked four questions on whether they think their partner would be envious of them (e.g., "My partner probably feels envious of me"). Contingent on sufficient internal consistency, these items will be summed and used to test whether **H2)** `r hyp2_min[1]` versus **H2~0~)** no significant indirect effect through anticipated partner envy from inequality to security consumption. This procedure has a notable strength by controlling the participant's income across levels of inequality (as the partner's money is the only income that changes). Since stake size will be manipulated within-subjects, two sets of responses for game-related self-report items and security consumption decisions will be analyzed using a mixed ANOVA model and mediation analyses.

#### A priori power analysis (Study 1a)

Using mean and standard deviation values of security expenditure from the pretest (*M*=`r round(base_mean,2)`, *SD*=`r round(base_sd,2)`), an a priori power analysis has been conducted using r package SuperPower[@lakens2021a] (see Figures \@ref(fig:mixed-design) and \@ref(fig:powerplot2) and and Tables \@ref(tab:2b2w-anova-exact) and \@ref(tab:2b2w-pairwise-power)). Given the unknown size of the hypothesized effect, this study will target an effect size of cohen's f = `r round(security_exact_5$main_results$cohen_f[1],2)`, which corresponds to a pattern of means yielding a cohen's d = .2 in an independent-samples t-test; a small effect according to statistical convention.[@cohen1988] To achieve 95% power with one tail, studies must have a minimum of n = `r round(cell_n_5,0)` participants per between-subjects cell, or a study-wide total of n = `r format(round(cell_n_5*2,0),big.mark = ",", scientific = FALSE)`.

### Study 1b: Inequality with real payoffs

The use of hypothetical scenarios in Study 1a may limit the ecological validity of any results. For instance, compared to hypothetical rewards, participants take fewer risks when incentivized by real payoffs and when they experience negative feedback.[@xu2016; @hertwig2001] Study 1b will serve to replicate Study 1a, with the sole difference being that **H1)** and **H1a)** are being tested when participants believe they are playing with a real partner, and that their decision could have real payoffs. Specifically, participants will be entered in a lottery where three participants will be selected to receive the payoff from one of their decisions.

This study will also use a mixed (2b\*2w) experimental design, manipulating inequality (yes vs. no) between-subjects, and the starting probabilities of successful thefts (50% vs. 60%) within-subjects. All conditions will have a stake size of \$300 (\$3 security cost and \$150 gains/losses from thefts). Study 1b will determine its targeted sample size by seeking 95% power for the observed main effect of inequality in Study 1a. If Study 1a fails to find a significant effect of inequality, Study 1b will again target an effect size of cohen's f = `r round(security_exact_5$main_results$cohen_f[1],2)` (n=`r format(round(cell_n_5*2,0),big.mark = ",", scientific = FALSE)`). Results will be analyzed using a mixed ANOVA and mediation analyses.

### Study 1c: Meritocracy

Groups will always be characterized by inequality through some type of currency, whether wealth, income, or status. Indeed, a popular idea is that inequality propels individuals towards productive behaviours.[@stearns1999] In this line of reasoning, the presence of inequality may be less of an issue compared to how the inequality was produced. The next experiment will test whether **H3)** `r tolower(hypotheses[2])` (Table \@ref(tab:design-table); versus **H3~0~)** no significant inequality\*merit interaction). As planned comparisons, Study 1c will examine whether **H3a)** hi-inequality/random income condition will have significantly higher rates of security consumption than all other conditions (i.e., **.1)** hi-inequality/merit income, **.2)** no-inequality/random income, & **.3)** no-inequality/merit income. Study 1c again varies inequality (yes vs. no) as the previous studies, but also manipulates whether one's income is seemingly allocated *randomly or based on task performance* (random vs. merit).

For this random vs. merit manipulation, all participants will complete a slider task, where they have one minute to correctly place 40 sliders at the midpoint of various lines that appear on their screens. Participants will then either be told that their incomes were allocated randomly, or because they correctly placed more sliders than their partner. The slider task has been used in several studies to examine the effects of effort in distributional preferences, and has demonstrated that individuals that win their income through performance are less likely to support redistribution of effort-driven income allocations.[@cassar2019]

As a manipulation check, before making their consumption decision, participants will be asked to report whether they believed the incomes were disbursed fairly (e.g.,"The way incomes were given for this game is fair"). Participants will be excluded from analyses if they complete fewer than 3 sliders, which would both indicate low engagement in the study and likely undermine the credibility of the manipulation. Results will be analyzed using a 2b\*2b between-subjects ANOVA.

<!--# competing hypothesis: deserved superiority will lead to a greater desire to retain relative positioning –> more consumption - could nullify or reverse -->

#### A priori power analysis (Study 1c)

Given the unknown size of the hypothesized inequality\*merit interaction, this study will target an effect size of cohen's f = .1. To achieve 95% power, studies must have a minimum of `r cell_n_2b2b_f01` participants per cell, or a study-wide total of n=`r format(cell_n_2b2b_f01*4,big.mark = ",", scientific = FALSE)`. This large sample size is necessary given the possibility of a "no-way" interaction, where only one condition is significantly different from the remaining conditions.[@simonsohn2015] In a no-way interaction, the simple effect of inequality is nullified depending on the moderator, and thus requires at least twice as many subjects per cell to achieve as much power as the simple effect. Figures \@ref(fig:2b2b-design) and \@ref(fig:2b2b-powerplot) display the pattern of means and per-condition sample size thresholds to obtain 95% power. Additionally, see Tables \@ref(tab:2b2b-anova-power) and \@ref(tab:2b2b-pairwise-power) for the power and effect sizes of the ANOVA and pairwise comparisons, respectively.

### Study 1d: Group membership

Fairness appears to have its roots in avoiding perceptions of partiality. For instance, young children will discard a resource in order to avoid sharing it unequally.[@shaw2012] Since Study 1c directly compares meritocratic versus random allocation of resources, it does not answer the more fundamental question: do people become distrustful and consume more security goods when inequality appears to be more discriminatory and unfair? Study 1d will keep a constant presence of inequality, but manipulate whether the partner is a member of an ingroup or outgroup, and whether the incomes are distributed randomly, or by a separate *ingroup individual*. Study 1d will use a 2b\*2b between subjects ANOVA to test whether **H4)** `r tolower(hypotheses[3])` (versus **H4~0~)** no significant outgroup\*agentic-distribution interaction). As planned comparisons, Study 1d will examine whether **H4a)** the outgroup partner/agentic-distribution condition will have significantly higher rates of security consumption than all other conditions (i.e., **.1)** outgroup/random, **.2)** ingroup/agentic, & **.3)** ingroup/random). Participants will complete the same fairness manipulation check as and be powered for the same effect as in Study 1c.


\newpage

# Data availability statement

Raw data can be accessed at [https://github.com/DallasNovakowski/inequality_security](https://github.com/DallasNovakowski/security_game){.uri}. 

# Code availability statement

Interactive demos of experimental games can be accessed at <br> <https://security-g.herokuapp.com/demo>, and experimental code can be accessed at <https://github.com/DallasNovakowski/security_game>. Preprocessing and analysis code is available at <https://github.com/DallasNovakowski/inequality_security>.

\newpage

# References

\begingroup

::: {#refs custom-style="Bibliography"}
:::

\endgroup

\newpage

# Acknowledgements

The funders had no role in study design, data collection and analysis, decision to publish or preparation of the manuscript.

The manuscript, including the figures and statistical analyses, was created using R (Version 4.1.1, R Core Team, 2021), RStudio (Version 1.4.1717, RStudio team, 2021), with major contributions from R-packages *tidyverse* (Version 1.3.1; Wickham et al., 2019), *lme4* (Version 1.1.27.1; Bates, Mächler, Bolker, & Walker, 2015), *afex* (Version 1.0, Singmann, Bolker, Westfall, Aust & Ben-Shachar, 2021),  *bookdown* (Version 0.24; Xie, 2021) and *Superpower* (Version 0.1.2; Lakens & Caldwell, 2021). Other useful packages can be found listed in the supplementary material, and in the shared code scripts.

```{r create_r-references, eval = F}
# r_refs(file = "r-references.bib")
ayy <- papaja::cite_r(file = "r-references.bib")
```

# Author contributions

All authors edited the manuscript and approved the final version. D.N. designed the studies and wrote the manuscript, in consultation with M.M.. D.N. collected, processed, and analysed the data. The study designs were based on theory developed by D.N.

# Competing interests

The authors declare no competing interests.

\newpage

<!-- # Figures & Figure captions -->

<!-- \newpage -->

# Tables

```{r design-table}
design <- design[-5,]

kd <- knitr::kable(design, caption = 'Design table', format = "latex",
                   # booktabs = TRUE, 
                   col.names = design_columns) %>%
  kable_styling(full_width = FALSE) %>%
  column_spec(1,width = "1.7in") %>%
  column_spec(2,width = "1.3in") %>%
  column_spec(3,width = "1.7in") %>%
  column_spec(4,width = "1.3in") %>%
  column_spec(5,width = "1.7in") %>%
  kableExtra::landscape()

kd
```


<!-- \clearpage -->

<!-- # (APPENDIX) Appendix {-}  -->

<!-- # Appendix A -->

\clearpage

# Supplementary material {.unnumbered}

\beginsupplement

```{r, child=c('supp_materials.Rmd')}

```
