---
title: "ICVS 1999 Prep - Screening and Cleaning"
author: "Dallas Novakowski"
date: "19/05/2022"
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(reshape)

load(file = "C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/data_1999.RData")

source(here::here("scripts", "s2_01_icvs_scripts.r"), local = knitr::knit_global())

load(file = here::here("output","icvs_helper.RData"))

potential_ivs <- gsub(x = potential_ivs, pattern = "_5_years", replacement = "_5yrs")  

```

```{r ncols and nrow, warning=F, message=FALSE, include = TRUE}
sweep4_nrow <- nrow(data_1999)
sweep4_nrow

ncol(data_1999)

summary(data_1999$country)[summary(data_1999$country)>0]


sweep4_countries <- sum(summary(data_1999$country)>0)
sweep4_countries

summary(data_1999[,reduced_var[[1]]])
```

```{r questionnaires, warning=F, message=FALSE}
kable_summary(data_1999$questionnaire_used)
```


```{r responders}
responders_1999 <- filter(data_1999, is.na(prev_refusal) & (prev_do_not_know == 0 | is.na(prev_do_not_know)))
nrow(responders_1999)
sum(summary(responders_1999$country)>0)
```


```{r questionnaires-after-exclusions}
kable_summary(responders_1999$questionnaire_used)
kable_summary(responders_1999$questionnaire_based_on)
```


```{r questionnaire-missingness}
questionnaire_screen <- responders_1999 %>% group_by(questionnaire_used) %>% 
  dplyr::count(country)

# view number of responses for each country*sweep
questionnaire_screen_c <- cast(questionnaire_screen,country~questionnaire_used)

# compute total number of participants
questionnaire_screen_c$total <- rowSums(questionnaire_screen_c[,2:ncol(questionnaire_screen_c)], na.rm=TRUE)

questionnaire_screen_c$total[questionnaire_screen_c$total == 0] <- NA

questionnaire_screen_c
```


```{r missing country}
na_tally(responders_1999,country,reduced_var[[4]]) %>%
    dplyr::bind_rows(dplyr::summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.logical), sum),
                      across(where(is.character), ~"Total")))
```

```{r ultra min prevention variables}
responders_1999 <- filter(responders_1999, country != "Switzerland")

#caretaker missing 7 (after removing switzerland)
# insurance missing 16
# arrangment missing 9
# lights on missing 45

# fence missing 2 (swaziland; lesotho)
# watch scheme missing 2 (mozambique, Japan)
# Gun ownership missing 1 (Estonia)

prevention_min_1999 <- c("prev_burglar_alarm", "prev_special_door_locks", "prev_special_grills", "prev_a_watch_dog", "prev_high_fence", "prev_watch_scheme","gun_ownership")
```


```{r rowsums security, warning=FALSE, message = F}
responders_1999[ , prevention_min_1999] <- sapply(responders_1999[ , prevention_min_1999],
                                             function(x) as.numeric(as.character(x)))

# create total security variable
responders_1999$total_security <- rowSums(responders_1999[ , prevention_min_1999], na.rm = TRUE) * NA ^ (rowSums(!is.na(responders_1999[ , prevention_min_1999])) == 0)

# any NA values are coded as 0
responders_1999$total_security[is.na(responders_1999$total_security)] <- 0

summary(responders_1999[ , prevention_min_1999])

summary(responders_1999$total_security)

hist_plot(responders_1999,total_security) + xlab("Reported security measures, Sweep 4")

```

```{r after exclusions}
summary(responders_1999$country)[0!= summary(responders_1999$country)]
```

Mozambique only has 3 responses

```{r ivna}
responders_1999 <- filter(responders_1999, country != "Mozambique")

colnames(responders_1999) <- gsub(x = colnames(responders_1999), pattern = "_5_years", replacement = "_5yrs")  


# count nonmissing
nrow(responders_1999) - colSums(is.na(responders_1999[,potential_ivs]))



# potential_ivs <- gsub(x = potential_ivs, pattern = "_5_years", replacement = "_5yrs")  

# array of variables with more than 40,000 responses
var_40k <- (nrow(responders_1999) - colSums(is.na(responders_1999[,potential_ivs])))[(nrow(responders_1999) - colSums(is.na(responders_1999[,potential_ivs]))) > 40000]
 
responders_1999[,names(var_40k)] %>%     
 summary()
```

```{r victimization-variables, warning=F, message=F}
# create small dataset with victimization items recoded
# colnames(victim_1999) <- gsub(x = colnames(victim_1999), pattern = "_5_years", replacement = "_5yrs")  
# 
# 
victim_1999 <- responders_1999[,names(var_40k)] %>%
 dplyr::select(ends_with("_5yrs")) %>%
  mutate_all(funs(dplyr::recode(., "no owner" = 0, "yes" = 1, "no" = 0, "do not know" = NA_real_, `-1`=0, `1`=1, `2` = 0, `0` = 0,  .default = NA_real_)))

vic_var_drop <- names(select(victim_1999,-c("assault_5yrs","sexoff_5yrs"
                                            ,"motortheft_5yrs"
                                            )))

responders_1999[,vic_var_drop] <- responders_1999[,vic_var_drop] %>%
  mutate_all(funs(dplyr::recode(., "no owner" = 0, "yes" = 1, "no" = 0, "do not know" = NA_real_, `-1`=0, `1`=1, `2` = 0, `0` = 0,  .default = NA_real_)))

# 
# #change victimization column names
# 
# 
# #replace victimization variables with updated ones
# responders_1999[,colnames(victim_1999)] <- victim_1999
# 
# # missingness tally
# na_tally(responders_1999,country,names(victim_1999)) %>%
#       bind_rows(dplyr::summarise(.,
#                       across(where(is.numeric), sum),
#                       across(where(is.logical), sum),
#                       across(where(is.character), ~"Total")))

```

```{r sum victim, echo=F, message=F}


vic_var_assault <- names(select(victim_1999,-c("sexoff_5yrs"
                                               ,"motortheft_5yrs"
                                               )))

# currently keeping
summary(victim_1999[,vic_var_drop])



# filtering to remove missing variables for summing
responders_1999 <-
  filter(responders_1999,
         (
           !is.na(cartheft_5yrs) &
             !is.na(bicyctheft_5yrs) &
             !is.na(burglar_5yrs) &
             !is.na(attempt_5yrs) &
             !is.na(robbery_5yrs) &
             !is.na(pers_theft_5yrs)
         ))

summary(responders_1999[,vic_var_drop])
nrow(responders_1999)

responders_1999$num_victim_5yr <- rowSums(responders_1999[,vic_var_drop])


summary(responders_1999$num_victim_5yr)


hist_plot(responders_1999,num_victim_5yr) + xlab("Reported victimizations, Sweep 4")

moments::skewness(responders_1999$num_victim_5yr)
```



```{r transform more predictor variables}
names(var_40k[-which(names(var_40k) %in% names(victim_1999))])

#make analysis variables
responders_1999 <- responders_1999 %>%
  dplyr::mutate(male = ifelse(gender=="male", 1, 0),
        age_num = dplyr::recode(as.numeric(age),`1` = 19, `2`=	24, `3` =	29,			
                        `4` =	34, `5` = 	39,`6` =	44,`7` =	49,`8` =	54,
                        `9` =	59,`10` =	64, `11` =	69, `12` = 75, `13` = NA_real_), 
            immigrant = ifelse(immigrant_status=="yes, self", 1, 0),
            household_size_num = dplyr::recode(as.numeric(household_size), `11` = NA_real_),
        
        
#housing - apartment as reference?
        apartment = ifelse(type_of_house == "flat appartment", 1, ifelse(type_of_house =="(semi)detached" | type_of_house =="terrace/row house", 0, ifelse(type_of_house == "institution" | type_of_house == "boat caravan other", 0, ifelse(type_of_house == "shanties", 0, NA)))),
        house = ifelse(type_of_house =="(semi)detached" | type_of_house =="terrace/row house", 1, ifelse(type_of_house == "flat appartment", 0, ifelse(type_of_house == "institution" | type_of_house == "boat caravan other", 0, ifelse(type_of_house == "shanties", 0, NA)))),         
        institution = ifelse(type_of_house =="(semi)detached" | type_of_house =="terrace/row house"| type_of_house == "boat caravan other", 0, ifelse(type_of_house == "institution", 1, ifelse(type_of_house == "flat appartment", 0, ifelse(type_of_house == "shanties", 0, NA)))),
        shanty = ifelse(type_of_house =="(semi)detached" | type_of_house =="terrace/row house"| type_of_house == "boat caravan other", 0, ifelse(type_of_house == "flat appartment", 0, ifelse(type_of_house == "institution", 0, ifelse(type_of_house == "shanties", 1, NA)))),
        other_house = ifelse(type_of_house == "boat caravan other", 1, ifelse(type_of_house == "flat appartment" | type_of_house =="(semi)detached" | type_of_house =="terrace/row house", 0, ifelse(type_of_house == "institution", 0, ifelse(type_of_house == "shanties", 0, NA)))),
      
        employed = ifelse(occupation == "looking for work" | occupation == "keeping home" | occupation == "retired, disabled" | occupation == "still at school" | occupation == "other", 0, ifelse(occupation == "working" | occupation == "army", 1, NA)),

        income_quartile = ifelse(income == "upper 25%", 4, ifelse(income == "50-75%", 3, ifelse(income == "25-50%", 2, ifelse(income == "lower 25%", 1,NA)))),
        income_tophalf  = ifelse(income == "upper 25%" | income == "50-75%" | income == "upper 50% (unknown)", 1, ifelse(income == "lower 50% (unknown)" | income == "lower 25%" | income == "25%-50%",0,NA)),


        years_edu = as.numeric(as.character(years_of_education)),

        city_size_k = as.numeric(ifelse(town_size == "-10,000", 0, ifelse(town_size == "10,000-50,000", 10,ifelse(town_size == "50,000-100,000",50, ifelse(town_size == "100,000-500,000",100,ifelse(town_size == "500,000-1,000,000",500,ifelse(town_size == "1,000,000+",1000,NA))))))),

        partnered = ifelse(marital_status == "married" | marital_status == "living together", 1, ifelse(marital_status == "single" | marital_status == "divorced" | marital_status == "widowed", 0, NA)),

        police_effective = ifelse(police_good_job_4_cat == "a very poor job",1,ifelse(police_good_job_4_cat == "a fairly poor job",2, ifelse(police_good_job_4_cat == "a fairly good job",3, ifelse(police_good_job_4_cat == "a very good job",4, NA)))),

        feel_safe_dark = ifelse(feel_safe_after_dark == "very unsafe",1,ifelse(feel_safe_after_dark == "bit unsafe",2, ifelse(feel_safe_after_dark == "fairly safe",3, ifelse(feel_safe_after_dark == "very safe",4, NA)))),

        break_in_threat = ifelse(break_in_chances == "not very likely",1,ifelse(break_in_chances == "likely",2, ifelse(break_in_chances == "very likely",3, NA))),

        outing_freq = ifelse(going_out == "never",1,ifelse(going_out == "less",2, ifelse(going_out == "once a month",3,ifelse(going_out == "once a week",4, ifelse(going_out == "almost every day",5, NA))))))

responders_1999$age_num <- responders_1999$age_num -2.5
responders_1999 <- filter(responders_1999, institution != 1 | institution == 0 | is.na(institution))
```

```{r demo-filter}
responders_1999 <- filter(responders_1999, !is.na(age_num) & !is.na(male) & !is.na(num_victim_5yr))

nrow(responders_1999)
sum(summary(responders_1999$country)>0)

responders_1999$male <- factor(responders_1999$male)
responders_1999$employed <- factor(responders_1999$employed)
responders_1999$partnered <- factor(responders_1999$partnered)
```


```{r}
new_iv <- names(responders_1999[,ncol(responders_1999):sum(ncol(responders_1999),-18)])

na_tally(responders_1999,country,all_of(new_iv)) %>%
    bind_rows(dplyr::summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.logical), sum),
                      across(where(is.character), ~"Total")))
```

Hungary missing housing items
Slovenia missing employed and education
Japan missing household size

Mozambique missing education
south africa, botswana, lesotho, and swaziland missing income 



```{r drop-hungary}
responders_1999 <- filter(responders_1999, country != "Hungary" & !is.na(house) & !is.na(partnered))

responders_1999[,c("age_num", "male", "employed", "partnered", "police_effective","years_edu","other_house","shanty","house","apartment","household_size_num", "income_quartile", "income_tophalf")] %>%
 summary()
```

```{r}
responders_1999[,names(var_40k)] %>%     
 summary()
```

minimal: 

moderate missingness (~1000): going out, break in chances, feeling safe after dark

greater missingness (1000-5000):  persons over 16

substantial missingness (5,000 10,000): years of education, occupation

```{r}
winz_function <- function(df) {
  df$age_cent <- df$age_num - mean(df$age_num)
  
  df$security_winz <- DescTools::Winsorize(df$total_security)
  df$security_0_1 <- as.factor(dplyr::recode(df$security_winz, `0` = 0L,
                           `1` = 1L,.default = NA_integer_))
  df$security_1_2 <- as.factor(dplyr::recode(df$security_winz, `1` = 0L,
                           `2` = 1L,.default = NA_integer_))
  df$security_2_3 <- as.factor(dplyr::recode(df$security_winz, `2` = 0L,
                           `3` = 1L,.default = NA_integer_))

  df$num_victim_5yr_winz <- DescTools::Winsorize(df$num_victim_5yr)
  
  df$num_victim_5yr_0_1 <- as.factor(dplyr::recode(df$num_victim_5yr_winz, `0` = 0L,
                           `1` = 1L,.default = NA_integer_))
  df$num_victim_5yr_1_2 <- as.factor(dplyr::recode(df$num_victim_5yr_winz, `1` = 0L,
                           `2` = 1L,.default = NA_integer_))
  
  df$police_eff_cent <- df$police_effective - mean(df$police_effective, na.rm = T)
  df$income_cent <- df$income_quartile - mean(df$income_quartile, na.rm = T)
  df$safe_dark_cent <- df$feel_safe_dark - mean(df$feel_safe_dark, na.rm = T)
  df$break_in_threat_cent <- df$break_in_threat - mean(df$break_in_threat, na.rm = T)
  df
}

responders_1999<- winz_function(responders_1999)

# create dataframe averaging victimization within each country
cdata <- plyr::ddply(responders_1999, c("country"), summarise,
               victim_nation_mean = mean(num_victim_5yr),
               percent_house = mean(house))

cdata$victim_nation_winz <- DescTools::Winsorize(cdata$victim_nation_mean)
cdata$victim_nation_cent <- cdata$victim_nation_mean - mean(cdata$victim_nation_mean)
cdata$victim_nation_wc <- cdata$victim_nation_winz - mean(cdata$victim_nation_winz)



# append full dataset with country-level means of victimizatoin
responders_1999 <- cdata %>%
  left_join(responders_1999, by = c("country"))
```

`ordered(security_total) ~ gini_1999_03 + gdppc_1999_03 + nation_victim_5yr + num_victim_5yr + male + age_num + house + other_house + shanty + partnered + (1|country)`

housing no good as a national variable; not randomly sampled (e.g., some taken from just one, heavily urbanized city; unless you believe that the country of georgia has 95% of people living in apartments)

```{r save-data}
responders_1999_countries <- names(summary(responders_1999$country)[summary(responders_1999$country)>0])
responders_1999_n <- nrow(responders_1999)

responders_1999_countries
responders_1999_n

cgwtools::resave(responders_1999,responders_1999_countries,responders_1999_n,file = "C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/data_1999.RData")
```

