---
title: "ICVS exploration"
date: "`r Sys.Date()`"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_float: true
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) # pipe
library(purrr) # map function
library(car) #qqPlot
library(survey)
# library(nloptr) # for bobyqa used in svy2lme

load("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/iv_2005.RData")

load("C:/Users/dalla/Google Drive/offline_data_files/icvs_pwt_swiid/data/icvs_joined_samples.RData")

load(file = here::here("output","icvs_output.RData"))

source(here::here("scripts", "s2_01_icvs_scripts.r"), local = knitr::knit_global())
# source(here::here("scripts", "s2_03_icvs_sample_models.r"), local = knitr::knit_global())
```

# Analysis procedure

The following describes two survey sweeps of the International Crime Victimization Survey conducted across 2004-6. At first, this data had `r sweep5_nrow` respondents and `r sweep5_countries` countries. Analyzing archival data, the so-called "garden of forking paths [\@gelman2014] allows a researcher to pick and choose those data and analyses that best suit their preferred story [\@simmons2011]. As such, transparently showcasing the different analysis paths is an important measure to help reduce the encroachment of model selection bias.

As mentioned in the manuscript, including certain variables (e.g., perceived police effectiveness) requires dropping certain countries of Japan, Peru, and Iceland (`mod_joined1`: k = `r sum(summary(as.factor(mod_joined1$country))>0)`, and n = `r format(nrow(mod_joined1), big.mark = ",", scientific = FALSE)`.

By comparison, excluding these three variables yields a sample of k = `r sum(summary(as.factor(iv_joined1$country))>0)` countries and a total n = `r format(nrow(iv_joined1), big.mark = ",", scientific = FALSE)` participants. The reported analyses will focus on the latter, larger sample specification, but this descriptives section (and supplementary analyses) examine whether a reduced model (i.e., excluding assaults, income quartile, and perceived police effectiveness) change between these sample subsets.

# Exploration

Starting with scatterplots. Nation Values with a cook's distance of 4 times the mean cook's distance values are flagged

```{r iv-descriptives}
# Fixed effect variables in model
summary(iv_joined1[,c("total_security","num_victim_5yr", "age_cent", "employed", "male")])

summary(gd[,c("gini_2004_6", "gdppc_2004_6")])
```

```{r security summary}
summary(iv_joined1$total_security)

normality_stats(iv_joined1$total_security)
```

## Larger (focal) dataset

To give a clearer picture of the data being included, data exploration will first be presented for the more complete dataset, then presented with the smaller subset.

### Influential observations

```{r influence screen, warning=F, message=FALSE, echo=FALSE}
gd$country <- as.character(gd$country)

victim_age <- ggplot(iv_joined1, aes(x = age_cent, y = num_victim_5yr)) +
    geom_point(alpha = .02, position = "jitter", color = "#40B0A6") + geom_smooth(alpha = .25, color = "red", span = 1.25) + geom_smooth(method = "lm", alpha = .25) + theme_minimal()+ ggtitle("Age & victimization")


gini_gdp_full <- gd %>% 
  mutate(cooks = cooks.distance(lm(gini_2004_6 ~ gdppc_2004_6)),
         outlier_country = ifelse(cooks > 4*mean(cooks), country, NA)) %>%
  ggplot(aes(x = gini_2004_6, y = gdppc_2004_6, label = outlier_country)) +
  geom_point(position = "jitter") + geom_smooth(alpha = .25, color = "red", span = 1.25) + geom_smooth(method = "lm", alpha = .25) + ggtitle("Gini & GDP - Full") + theme_minimal() + coord_cartesian(xlim = c(20, 65), ylim = c(-10000, 85000)) + ggrepel::geom_label_repel()


# victim GINI plot with outliers
victim_gini_full <- gd %>% 
  mutate(cooks = cooks.distance(lm(num_victim_5yr ~ gini_2004_6)),
         outlier_country = ifelse(cooks > 4*mean(cooks), country, NA)) %>%
  ggplot(aes(x = gini_2004_6, y = num_victim_5yr, label = outlier_country)) +
  geom_point(position = "jitter") + geom_smooth(alpha = .25, color = "red", span = 1.5) + geom_smooth(method = "lm", alpha = .25) + ggtitle("Victimization & Gini - Full") + theme_minimal() + coord_cartesian(xlim = c(20, 65), ylim = c(.25, 1.125)) + ggrepel::geom_label_repel()

ggsave(here::here("figures", "victim_viz.png"),
       height = 3)

# victim GDP plot with outliers
victim_gdp_full <- gd %>% 
  mutate(cooks = cooks.distance(lm(num_victim_5yr ~ gdppc_2004_6)),
         outlier_country = ifelse(cooks > 4*mean(cooks), country, NA)) %>%
  ggplot(aes(x = gdppc_2004_6, y = num_victim_5yr, label = outlier_country)) +
  geom_point(position = "jitter") + geom_smooth(alpha = .25, color = "red", span = 1.25) + geom_smooth(method = "lm", alpha = .25) + ggtitle("Victimization & GDP - Full") + theme_minimal() + coord_cartesian(xlim = c(0, 85000), ylim = c(.25, 1.125)) + ggrepel::geom_label_repel(label.size = 0.02)


gridExtra::grid.arrange(victim_age, victim_gini_full,victim_gdp_full, gini_gdp_full, nrow = 2, ncol=2)


```

```{r, include=FALSE}
# victim_gini_full
# 
# ggsave(here::here("figures", "victim_viz.png"),
#        height = 3)

```

### Model-based exploration

Now there are obviously some heavy tails in this data.

```{r estimation preview2, warning=T, message=FALSE}
# call basic unweighted model
summary(lme1)
```

Further seen in the normality statistics for the residuals - high skewness and kurtosis

```{r model skewness kurtosis,  warning=F, message=FALSE}
moments::skewness(resid(lme1))
moments::kurtosis(resid(lme1))
```

Indeed, this model's residuals are definitely not normally-distributed, with a skewness of `r round(moments::skewness(resid(lme1)),2)`, and a kurtosis of *`r round(moments::kurtosis(resid(lme1)),2)`.* However, this is not yet the time to examine the residual plots - we need to explore and treat any potential influential observations that might be exacerbating

> Note that this is based on simple aggregates. It is not based on empirical Bayesian estimates Doesn't take sampling variation due to sampling of into account Therefore a rough guide, but only rough

#### Cook's distance ( R package influence.ME )

```{r cooksd model, warning=F, message=FALSE, fig.show="hold", out.width="50%"}
# influential cases (Cook's Distance. Any values over 1 are likely to be significant outliers)
cooksD <- cooks.distance(lme1)
summary(cooksD)
```

```{r cooksd lv2, warning=F, message=FALSE, echo=FALSE, fig.show="hold", out.width="50%"}
# look for influential observations in nation-level
plot(cooksD, main="Model-wide Cook's D * BEWARE Y AXIS", ylim=c(0,.01))
abline(h = 4/nrow(iv_joined1), lty = 2, col = "steelblue") # add cutoff line
abline(h = 4*mean(cooksD, na.rm=T), col="red") 


# nation_cook <- cooks.distance(lm(gini_2004_6 ~ gdppc_2004_6, gd))
# 
# plot(nation_cook, main="Nation-level (Gini ~ GDP) Cook's D", ylim=c(0,1))
# abline(h = 4/nrow(gd), lty = 2, col = "steelblue") # add cutoff line
# abline(h = 4*mean(nation_cook, na.rm=T), col="red") 
# text(x=1:length(nation_cook)+1, y=nation_cook, labels=ifelse(nation_cook>4*mean(nation_cook, na.rm=T), gd$country,""), col="red")

victim_nation_cook <- cooks.distance(lm(num_victim_5yr ~ gini_2004_6 + gdppc_2004_6, gd))

plot(victim_nation_cook, main="Cook's D, Victimization ~ Gini + GDP", ylim=c(0,1))
abline(h = 4/nrow(gd), lty = 2, col = "steelblue") # add cutoff line
abline(h = 4*mean(victim_nation_cook, na.rm=T), col="red") 
text(x=1:length(victim_nation_cook)+1, y=victim_nation_cook, labels=ifelse(victim_nation_cook>4*mean(victim_nation_cook, na.rm=T), gd$country,""), col="red") 

 
 # lablist.x<-gd$country
# axis(1, at=seq(1, 31, by=1), labels = lablist.x)
# text(x = seq(0, 31, by=1),par("usr")[3] - 0.45, labels = lablist.x, srt = 45, pos = 2, xpd = FALSE)


 
 # add labels
# 
# axis(1, by=1, at = 1:31, labels = gd$country)

# 
# gd %>% 
#   mutate(cooks = cooks.distance(lm(gini_2004_6 ~ gdppc_2004_6)),
#          outlier_country = ifelse(cooks > 4/nrow(.), country, NA)) %>% 
#   ggplot(aes(x = gini_2004_6, y = gdppc_2004_6, label = outlier_country)) +
#   geom_point() + ggrepel::geom_label_repel()

```

A model-wide Cook's D shows that any given observation has a small influence on the entire model - to be expected with such a large sample. This is why it's important to break the analysis of influential variables into multiple levels

We can see that level-2 analysis yields much larger Cook's D values. The cooks distance values are all well under 1. Importantly, this is the same for the nation-level variables; they have a smaller sample size - and are thus the more susceptible to influential observations.

**However, there are a few observations at nation-level that exceed the 4/n threshold, and some that exceed 0.5. Generally, there are changes in the slope, but ultimately the direction doesn't change**

## Handling influential observations - Winzorization

Since these influential observations are not a mistake - but rather truly extreme data, we will not delete the data, but instead winzorize these extreme values.

```{r winzorize descriptives, warning=F, message=FALSE, echo=FALSE, fig.show="hold", out.width="50%"}


# iv_joined1[["num_victim_5yr_winz"]] <-  DescTools::Winsorize(iv_joined1[["num_victim_5yr"]])
# 
# 
# iv_joined1[["gdppc_2004_6_winz"]] <-  DescTools::Winsorize(iv_joined1[["gdppc_2004_6"]])
# 
# iv_joined1[["gdppc_2004_6_wc"]] <-  iv_joined1[["gdppc_2004_6_winz"]] - mean(iv_joined1[["gdppc_2004_6_winz"]])
# 
# iv_joined1[["gdppc_2004_6_scale"]] <- scale(iv_joined1[["gdppc_2004_6"]])
# 
# iv_joined1[["gdppc_2004_6_ws"]] <- scale(iv_joined1[["gdppc_2004_6_winz"]])
# 
# 
# iv_joined1[["security_winz"]] <-  DescTools::Winsorize(iv_joined1$total_security)
# 

iv_joined1[,c("num_victim_5yr", "num_victim_5yr_winz", "gdppc_2004_6", "gdppc_2004_6_winz", "gini_2004_6_cent", "gini_2004_6_winz")] %>%
  summary()

gd[,c("gdppc_2004_6", "gdppc_2004_6_winz", "gini_2004_6", "gini_2004_6_winz")] %>%
  summary()


```

Dramatic changes to the max values, less-so for the mean values (with the exception of number of victimizations)

```{r winzorize histograms, warning=F, message=FALSE, echo=FALSE}

gridExtra::grid.arrange(


hist_plot(iv_joined1,num_victim_5yr,"Regular number of victimizations"),
hist_plot(iv_joined1,num_victim_5yr_winz,"Winzorized number of victimizations"),

hist_plot(iv_joined1,total_security,"Regular security consumption"),
hist_plot(iv_joined1,security_winz,"Winzorized security consumption"),


hist_plot(gd,gdppc_2004_6,"Regular GDP"),
hist_plot(gd,gdppc_2004_6_winz,"Winzorized GDP"),

hist_plot(gd,gini_2004_6,"Regular Gini"),
hist_plot(gd,gini_2004_6_winz,"Winzorized Gini"),

nrow = 4, ncol=2)

```

```{r winz-plot-iv}
victim_gini_full <- gd %>% 
  mutate(cooks = cooks.distance(lm(num_victim_5yr ~ gini_2004_6_winz)),
         outlier_country = ifelse(cooks > 4*mean(cooks), country, NA)) %>%
  ggplot(aes(x = gini_2004_6_winz, y = num_victim_5yr, label = outlier_country)) +
  geom_point(position = "jitter") + geom_smooth(alpha = .25, color = "red", span = 1.5) + geom_smooth(method = "lm", alpha = .25) + ggtitle("Victimization & Winzorized Gini") + theme_minimal() + coord_cartesian(xlim = c(20, 50), ylim = c(.25, 1.125)) + ggrepel::geom_label_repel()

victim_gini_full
```


### Multicolinnearity

-   **Check**: VIF/tolerances
-   **Fix** remove values \>= 5

```{r VIF1, warning=F, message=FALSE, echo = F}
performance::check_collinearity(lme1)

#car::vif(lme1) # display VIF values
```

## Smaller dataset

```{r descriptives, echo = F}
# Fixed effect variables in model
summary(mod_joined1[,c("num_victim_5yr", "age_cent", "employed", "male")])
```

```{r mod-descriptives, echo=F}
summary(mod_joined1[,c("num_victim_5yr", "age_cent", "employed", "male")])

mod_joined1[,vic_var_assault] <- lapply(mod_joined1[,vic_var_assault], factor)

summary(mod_joined1[,vic_var_assault])

mod_joined1[,prevention_min2][is.na(mod_joined1[,prevention_min2])] <- 0
mod_joined1[,prevention_min2] <- lapply(mod_joined1[,prevention_min2], factor)

summary(mod_joined1[,prevention_min2]) %>%
  kbl(booktabs = T) %>%
  kable_styling(latex_options = c("striped"))

summary(gd_mod[,c("gini_2004_6", "gdppc_2004_6")])
```

### Influential observations

```{r mod-influence screen, warning=F, message=FALSE, echo=FALSE}

gd_mod$country <- as.character(gd_mod$country)


victim_age <- ggplot(mod_joined1, aes(x = age_cent, y = num_victim_5yr)) +
    geom_point(alpha = .02, position = "jitter", color = "#40B0A6") + geom_smooth(alpha = .25, color = "red", span = 1.25) + geom_smooth(method = "lm", alpha = .25) + theme_minimal()+ ggtitle("Age & victimization")


gini_gdp_full <- gd_mod %>% 
  mutate(cooks = cooks.distance(lm(gini_2004_6 ~ gdppc_2004_6)),
         outlier_country = ifelse(cooks > 4*mean(cooks), country, NA)) %>%
  ggplot(aes(x = gini_2004_6, y = gdppc_2004_6, label = outlier_country)) +
  geom_point(position = "jitter") + geom_smooth(alpha = .25, color = "red", span = 1.25) + geom_smooth(method = "lm", alpha = .25) + ggtitle("Gini & GDP - Full") + theme_minimal() + coord_cartesian(xlim = c(20, 65), ylim = c(-10000, 85000)) + ggrepel::geom_label_repel()


# victim GINI plot with outliers
victim_gini_full <- gd_mod %>% 
  mutate(cooks = cooks.distance(lm(num_victim_5yr ~ gini_2004_6)),
         outlier_country = ifelse(cooks > 4*mean(cooks), country, NA)) %>%
  ggplot(aes(x = gini_2004_6, y = num_victim_5yr, label = outlier_country)) +
  geom_point(position = "jitter") + geom_smooth(alpha = .25, color = "red", span = 1.5) + geom_smooth(method = "lm", alpha = .25) + ggtitle("Victimization & Gini - Full") + theme_minimal() + coord_cartesian(xlim = c(20, 65), ylim = c(.25, 1.125)) + ggrepel::geom_label_repel()


# victim GDP plot with outliers
victim_gdp_full <- gd_mod %>% 
  mutate(cooks = cooks.distance(lm(num_victim_5yr ~ gdppc_2004_6)),
         outlier_country = ifelse(cooks > 4*mean(cooks), country, NA)) %>%
  ggplot(aes(x = gdppc_2004_6, y = num_victim_5yr, label = outlier_country)) +
  geom_point(position = "jitter") + geom_smooth(alpha = .25, color = "red", span = 1.25) + geom_smooth(method = "lm", alpha = .25) + ggtitle("Victimization & GDP - Full") + theme_minimal() + coord_cartesian(xlim = c(0, 85000), ylim = c(.25, 1.125)) + ggrepel::geom_label_repel(label.size = 0.02)



 
# victim_gini_full <- ggplot(gd, aes(x = gini_2004_6, y = num_victim_5yr)) +
#     geom_point() + geom_smooth(alpha = .25, color = "red", span = 1.25) + geom_smooth(method = "lm", alpha = .25) + 
#     theme_minimal() + ggtitle("Victimization & Gini - Full")  + coord_cartesian(xlim = c(20, 65), ylim = c(.25, 1.25))

# 
# 
# victim_gdp_full <- ggplot(gd, aes(x = gdppc_2004_6, y = num_victim_5yr)) +
#     geom_point() + geom_smooth(alpha = .25, color = "red", span = 1.25) + geom_smooth(method = "lm", alpha = .25) + 
#     theme_minimal() + ggtitle("Victimization & GDP - Full")  + coord_cartesian(xlim = c(1500, 85000), ylim = c(.25, 1.25))

gridExtra::grid.arrange(victim_age, victim_gini_full,victim_gdp_full, gini_gdp_full, nrow = 2, ncol=2)

```


### Model-based exploration

Now there are obviously some heavy tails in this data.

```{r basic-mod-summ}
# call basic unweighted model
summary(lme_mod1)
```

Further seen in the normality statistics for the residuals - high skewness and kurtosis

```{r mod-model skewness kurtosis,  warning=F, message=FALSE}
moments::skewness(resid(lme_mod1))
moments::kurtosis(resid(lme_mod1))
```

#### Cook's distance ( R package influence.ME )

```{r mod-cooksd model, warning=F, message=FALSE, fig.show="hold", out.width="50%"}
# influential cases (Cook's Distance. Any values over 1 are likely to be significant outliers)
cooksD <- cooks.distance(lme_mod1)
summary(cooksD)
```

```{r mod-cooksd lv2, warning=F, message=FALSE, echo=FALSE, fig.show="hold", out.width="50%"}
# look for influential observations in nation-level
plot(cooksD, main="Model-wide Cook's D * BEWARE Y AXIS", ylim=c(0,.01))
abline(h = 4/nrow(mod_joined1), lty = 2, col = "steelblue") # add cutoff line
abline(h = 4*mean(cooksD, na.rm=T), col="red") 


# nation_cook <- cooks.distance(lm(gini_2004_6 ~ gdppc_2004_6, gd))
# 
# plot(nation_cook, main="Nation-level (Gini ~ GDP) Cook's D", ylim=c(0,1))
# abline(h = 4/nrow(gd), lty = 2, col = "steelblue") # add cutoff line
# abline(h = 4*mean(nation_cook, na.rm=T), col="red") 
# text(x=1:length(nation_cook)+1, y=nation_cook, labels=ifelse(nation_cook>4*mean(nation_cook, na.rm=T), gd$country,""), col="red")

victim_nation_cook <- cooks.distance(lm(num_victim_5yr ~ gini_2004_6 + gdppc_2004_6, gd_mod))

plot(victim_nation_cook, main="Cook's D, Victimization ~ Gini + GDP", ylim=c(0,1))
abline(h = 4/nrow(gd_mod), lty = 2, col = "steelblue") # add cutoff line
abline(h = 4*mean(victim_nation_cook, na.rm=T), col="red") 
text(x=1:length(victim_nation_cook)+1, y=victim_nation_cook, labels=ifelse(victim_nation_cook>4*mean(victim_nation_cook, na.rm=T), gd_mod$country,""), col="red") 

 
 # lablist.x<-gd$country
# axis(1, at=seq(1, 31, by=1), labels = lablist.x)
# text(x = seq(0, 31, by=1),par("usr")[3] - 0.45, labels = lablist.x, srt = 45, pos = 2, xpd = FALSE)


 
 # add labels
# 
# axis(1, by=1, at = 1:31, labels = gd$country)

# 
# gd %>% 
#   mutate(cooks = cooks.distance(lm(gini_2004_6 ~ gdppc_2004_6)),
#          outlier_country = ifelse(cooks > 4/nrow(.), country, NA)) %>% 
#   ggplot(aes(x = gini_2004_6, y = gdppc_2004_6, label = outlier_country)) +
#   geom_point() + ggrepel::geom_label_repel()

```

## Handling influential observations - Winzorization

Since these influential observations are not a mistake - but rather truly extreme data, we will not delete the data, but instead winzorize these extreme values.

```{r mod-winzorize descriptives, warning=F, message=FALSE, echo=FALSE, fig.show="hold", out.width="50%"}

# mod_joined1[["num_victim_5yr_winz"]] <-  DescTools::Winsorize(mod_joined1[["num_victim_5yr"]])
# 
# mod_joined1[["gdppc_2004_6_winz"]] <-  DescTools::Winsorize(mod_joined1[["gdppc_2004_6"]])
# 
# mod_joined1[["gdppc_2004_6_wc"]] <-  mod_joined1[["gdppc_2004_6_winz"]] - mean(mod_joined1[["gdppc_2004_6_winz"]])
# 
# mod_joined1[["gdppc_2004_6_scale"]] <- scale(mod_joined1[["gdppc_2004_6"]])
# 
# mod_joined1[["gdppc_2004_6_ws"]] <- scale(mod_joined1[["gdppc_2004_6_winz"]])
# 
# 
# mod_joined1[["security_winz"]] <-  DescTools::Winsorize(mod_joined1$total_security)


mod_joined1[,c("num_victim_5yr", "num_victim_5yr_winz", "gdppc_2004_6", "gdppc_2004_6_winz", "gini_2004_6_cent", "gini_2004_6_winz")] %>%
  summary()
```

```{r mod-winzorize histograms, warning=F, message=FALSE, echo=FALSE}
gridExtra::grid.arrange(

hist_plot(mod_joined1,num_victim_5yr,"Regular number of victimizations"),
hist_plot(mod_joined1,num_victim_5yr_winz,"Winzorized number of victimizations"),

hist_plot(mod_joined1,total_security,"Histogram of security consumption"),
hist_plot(mod_joined1,security_winz,"Histogam of Winzorized security consumption"),


hist_plot(gd_mod,gdppc_2004_6,"Regular GDP"),
hist_plot(gd_mod,gdppc_2004_6_winz,"Winzorized GDP"),

hist_plot(gd_mod,gini_2004_6,"Regular Gini"),
hist_plot(gd_mod,gini_2004_6_winz,"Winzorized Gini"),

nrow = 4, ncol=2)

```

```{r hist-info, warning=F, message=FALSE, fig.show='hide', echo = F}
hist_plot(mod_joined1, total_security, "Histogram of Total Security")
ggsave(here::here("figures", "security_hist_mod.png"),
       height = 3)

hist_plot(mod_joined1, num_victim_5yr_assault, "Histogram of victimizations over past 5 years")
ggsave(here::here("figures", "victim_hist_mod.png"),
       height = 3)

hist_plot(iv_joined1,total_security,"Histogram of security consumption")
ggsave(here::here("figures", "security_hist_iv.png"),
       height = 3)

hist_plot(iv_joined1, num_victim_5yr, "Histogram of victimizations over past 5 years")
ggsave(here::here("figures", "victim_hist_iv.png"),
       height = 3)


```

### Multicolinnearity

-   **Fix** remove values \>= 5

```{r mod-VIF1, warning=F, message=FALSE}
performance::check_collinearity(lme_mod1)

#car::vif(lme_mod1) # display VIF values
```

# Limitations

The most notable weakness is that this is data from WEIRD countries. Other, less WEIRD countries (e.g., phillipines, nigeria, india) are better represented throughout sweeps 2 to 4

-   Unaccounted third variables (e.g., corruption)

Small clusters, limited representativeness.

# References
